### 2024-11-12 paper/main.tex
paper

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}

\title{Vlasov PINNs for L1-Lagrange Solar Wind modeling}
\author{Jorge D. Enciso}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

The solar wind phenomena is a vastly known event that arises from ionized solar outbursts into the stellar medium, commonly stimulating planets' magnetosphere. \cite{Gosling2007} The first record of a related event dates from 1859, popularly named "The Carrington event" after the English astronomer Richard Carrington (1826, 1875), who settled the intuition upon solar flares and geomagnetic fluctuations. Individuals from all around the glove sighted the Northern Lights, subtle variations in atmosphere's hues' tones, caused by ionized particles flowing through the ionosphere.

Several attempts to model Solar Wind's dynamics can be empirically pinpointed : analytical methods \cite{}, numerical modelling \cite{2006LNCS.3732..554E, 10.3389/fspas.2023.1105797, windmodelling1}, and machine learning approaches \cite{comp_1, comp_2, comp_3}. All of them with different purposes.

Traditional models and methodologies fail to account for the anysotropic nature of the Solar Wind. The Magnetohydrodynamics Model assumes a Boltzmann-Maxwellian distribution, doted with standard thermodynamical behavior. On the other hand, the Solar Wind is a collisionless plasma, precluding its inclusion solely within those boundaries. Thus, the present work emphasizes the usage of learnable neural operators engineered to analytically and empirically fit kinetic descriptions.

Moreover, kinetic frames based on Maxwellian electromagnetism tend to be computationally expensive for the microscopic and macroscopic frame:

\begin{equation}
    n_{\alpha} (\mathbf{r}, t) := \int f_{\alpha}(\mathbf{r}, \mathbf{v}, t) d^3v
\end{equation}

\begin{equation}
    \tau(r, t) := \sum q_{\alpha} n_{\alpha}
\end{equation}

\begin{equation}
    \mathbf{u_{\alpha}} (\mathbf{r}, t) := \frac{1}{n_{\alpha}(\mathbf{r}, t)} \int v f_{\alpha}(\mathbf{r}, v, t) d^3v
\end{equation}

\begin{equation}
    j(\mathbf{r}, t) := \sum q_{\alpha} n_{\alpha} \mathbf{u_{\alpha}}
\end{equation}

where $n_{\alpha}$ is the particle density, $u_{\alpha}$ the particle's drift velocity, $\tau(r, t)$ is the charge density, $j(r, t)$ is the current density, and $f_{\alpha}(r, v, t)$ is the Boltzmann's density function.

Yet modern physics informed machine learning research schemes enable faster inference time and less computational load: embedding systems' equations within the loss function and constraining network's backbone based on empirical approaches.

\section{Related Work}

\subsection{Solar Wind modelling}

\subsection{Physics informed Vlasov modeling}

\section{Kinetic Models}
In order to fully assume the magnetohydrodynamical model, interactions must converge to the Maxwell-Boltzmann thermodynamical distribution, implying a high colisionality degree, contrary to empirical observations of the Solar Wind. Therefore, the utilization of such a constrictive framework must be carefully evaluated before setting it up as a modeling constraints.

Moreover, trying to answer quasy-isotropic behavior on near-earth L1-Lagrange point, different research papers assume the bi-Maxwellian temperature distributions, carrying out the Maxwell-Boltzmann thermodynamical distribution. Nevertheless, this assumption seems overly ideal. As a possible option, the inclusion of numerical distributions could bring insights into the interactions between plasma ionized particles.

That's why, this work departs from the traditional approach, seeking the description of Solar Wind kinetic model without magnetohydrodynamical systems nor temperature distributions, fully describing these interactions with a physics informed data-driven structure as a substitution of empirical assumptions.

\subsection{Boltzmann's equation}
Within this frame, different kinetic statistical descriptors are suitable for the task as long as they make the correct assumptions. Therefore, the usage of Boltzmann's equation works perfectly within certain boundaries:

\begin{equation}
    \frac{\partial f}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha= \left(\frac{\delta f}{\delta t}\right)_{col}
\end{equation}

This result is widely known for describing the Maxwell-Boltzmann distribution under defined conditions, so an alternative interpretation that accounts for colisionless plasma should be integrated:

\subsection{Vlasov's equation}
The Vlasov's equation is the colisionless variant of the Boltzmann's equation. Therefore, it is a  suitable kinetic model for the Solar Wind, accounting for key properties of it.

\begin{equation}
    \frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{\mathbf{F}}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
\end{equation}

In this case, governed by electromagnetic interactions. Therefore, replacing the force $\mathbf{F}$ with the Lorentz Force:

\begin{equation}
    \frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
\end{equation}

\section{Physics Informed Machine Learning}
Usually, solely data-driven methods require huge volumes of high-quality data with black box algorithms that prevent interpretability, a key factor for physical modeling. The most recent research \cite{PINNS} advocate for physical informed virtual loss function terms, leading off to the vastly known field of Phyics Informed Neural Networks.

This models where created as numerical methods for solving Partial Differential Equations that embed physical modeling. It further extended to all fields that required PDE solving as a computationally efficient numerical alternative to other solvers.

\subsection{Approximating the density function with PINNs}
The most recent research, using PINNs, develop a way to not just learn functions, but to learn functional operators: Lagrangian, Hamiltonian, etc. This research \cite{} creates a neural network that takes the general coordinates $q$ and $\dot q$ as input, to provide the lagrangian as an output, and leverage the auto-differentiation capabilities of modern machine learning frameworks to enforce the euler-lagrange equation:

\begin{equation}
    \frac{d}{dt} \frac{\partial \mathcal{L}}{\partial \dot q} - \frac{\partial \mathcal{L}}{\partial q} = 0
\end{equation}

As usage of $q$ and $\dot q$ as inputs is very convenient for the auto-differentiation engines, making it further computationally feasible.

\section{Data}

\subsection{DSCOVR: Deep Space Climate Observatory}
DSCOVR, a joint mission between NASA and the National Oceanic and Atmospheric Administration, is a crucial observational platform for monitoring space weather \cite{nasa_dscovr}. Launched in 2015, DSCOVR's primary mission is to monitor and provide advanced warning of potentially hazardous space weather events such as solar flares and coronal mass ejections that could impact Earth.

It is equipped with two key instruments for measuring both energetic particle incidence and magnetic field parameters: the Faraday cup and the magnetometer from the PlasMag instrument \cite{nasa_dscovr}. The readings from these two sensors are crucial for virtually analyzing plasma dynamics near the L1 Lagrange point. These readings will be used as part of the core model data due to their real-time availability.

\subsection{ACE: Advanced Composition Explorer}
ACE, launched in 1997, provides continuous measurements of the solar wind and interstellar particles. It is equipped with several instruments designed to study the composition of solar and galactic particles, which are crucial for understanding the space weather environment. ACE's data helps in predicting geomagnetic storms and contributes to our understanding of the heliosphere.

\subsection{WIND}
The WIND spacecraft, launched in 1994, is part of the Global Geospace Science initiative. It provides comprehensive measurements of the solar wind, magnetic fields, and energetic particles. WIND's data is essential for understanding the fundamental processes of the solar wind and its interaction with the Earth's magnetosphere.

\section{Methods}

\subsection{Neural Architecture Search (NAS)}
As an intermediate step before training, a hyperparameter grid search algorithm based in Bayesian multi-objective optimization is employed to find the neural architecture. The Pareto front is plotted, selecting minimization of neural complexity and loss function.

\subsection{Loss function}
The neural network is trained using the standard gradient-descent based optimization algorithm: Adam. One Cycle learning optimizer is used to facilitate the learning rate plateau. Weight decay is implemented as a regularizer.

The loss function consists on a virtual term, three integral boundary condition terms, and a l2 regularizer term:

\begin{equation}
    \mathcal{L}_{virtual}\left(\mathbf{E}, \mathbf{B}, \mathbf{v}, \mathbf{r}; \mathbf{\theta}\right) = \left(\frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha \right)^2
\end{equation}

\begin{equation}
    \mathcal{L}_{BC_1} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} d^3\mathbf{r} - 1\right)^2
\end{equation}

\begin{equation}
    \mathcal{L}_{BC_2} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} - n_\alpha(\mathbf{r}, t) \right)^2
\end{equation}

\begin{equation}
    \mathcal{L}_{BC_3} \left(x; \mathbf{\theta}\right) = \left(\sum_{\alpha} q_\alpha n_\alpha  \mathbf{u}_\alpha- j(\mathbf{r}, t) \right)^2
\end{equation}

\begin{equation}
    \mathcal{L}_{regularizer} \left( \mathbf{\theta} \right) = \mathbb{E}_{w \in \mathbf{\theta}}\left[ w ^ 2 \right]
\end{equation}

Getting our final objective function:

\begin{equation}
    \mathcal{L}\left[\mathcal{L}_{virtual}, \mathcal{L}_{BC}, \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{virtual} + \alpha_2 \sum_{i} \mathcal{L}_{BC_i} + \lambda \mathcal{L}_{regularizer}
\end{equation}

\section{Results}

\section{Discussion}

\section{Conclusion}
This work demonstrated the effectiveness of Physics-Informed Neural Networks to model Solar Wind's behavior tied to diverse empirical and analytical formulations. By embedding the governing physical equations directly into the loss function, we can efficiently train a neural network to approximate the solution of this complex, non-linear system. Future work will focus on improving the kernel design and extending the approach to more sophisticated MHD scenarios, such as the interactions between Solar Wind and Sun's magnetosphere.

\bibliographystyle{plain}

\bibliography{references}
\end{document}

### 2024-11-12 paper/main.tex
default

@@
 
-\title{Vlasov PINNs for L1-Lagrange Solar Wind modeling}
+\title{Determining Solar Wind's distribution with Vlasov PINNs}
 \author{Jorge D. Enciso}
@@
 \begin{abstract}
+    %TODO
 \end{abstract}
@@
 
-Several attempts to model Solar Wind's dynamics can be empirically pinpointed : analytical methods \cite{}, numerical modelling \cite{2006LNCS.3732..554E, 10.3389/fspas.2023.1105797, windmodelling1}, and machine learning approaches \cite{comp_1, comp_2, comp_3}.
+Several attempts to model Solar Wind's dynamics can be empirically pinpointed : analytical methods \cite{BLUME202396}, numerical modelling \cite{10.3389/fspas.2023.1105797, windmodelling1, Gombosi_2018}, and machine learning approaches \cite{comp_2, comp_3, guastavino2024forecastinggeoffectiveeventssolar, sabbatini2023solarwindspeedestimate, https://doi.org/10.1029/2023SW003561}.
 All of them with different purposes.
@@
 
-\subsection{Solar Wind modelling}
+\subsection{Full compressible 3D MHD simulation of solar wind}
+On the matter of Solar Wind modeling, several numerical approaches are developed to understand near-Sun physics given certain boundary conditions.
+This research \cite{windmodelling1} creates a magnetohydrodynamical modeling scheme to understand these interactions under certain restrictions that approach heliospheric conditions.
+In this case, the current work is seeking more general descriptors of the Solar Wind, with real L1-Lagrange data as a cornerstone for the modeling scheme.
 
-\subsection{Physics informed Vlasov modeling}
+\subsection{Physics informed Neural Networks applied to the description of wave-particle resonance in kinetic simulations of fusion plasmas}
+
+\cite{kumar2023physicsinformedneuralnetworks}
+
+\subsection{Physics-Informed Neural Networks for Solar Wind Prediction}
+
+\cite{johnson2022physics}
 
 \section{Kinetic Models}
-In order to fully assume the magnetohydrodynamical model, interactions must converge to the Maxwell-Boltzmann thermodynamical distribution, implying a high colisionality degree, contrary to empirical observations of the Solar Wind.
+In order to fully assume the magnetohydrodynamical model, interactions must converge to the Maxwell-Boltzmann thermodynamical distribution for certain conditions, implying a high colisionality degree, contrary to empirical observations of the Solar Wind.
 Therefore, the utilization of such a constrictive framework must be carefully evaluated before setting it up as a modeling constraints.
 
-Moreover, trying to answer quasy-isotropic behavior on near-earth L1-Lagrange point, different research papers assume the bi-Maxwellian temperature distributions, carrying out the Maxwell-Boltzmann thermodynamical distribution.
+Moreover, trying to answer quasy-isotropic behavior on near-earth L1-Lagrange point, different research papers assume the bi-Maxwellian or Kappa temperature distributions \cite{Stansby_2018, Nicolaou_2018, Zouganelis_2004}, carrying out the Maxwell-Boltzmann thermodynamical distribution.
 Nevertheless, this assumption seems overly ideal.
@@
 
-That's why, this work departs from the traditional approach, seeking the description of Solar Wind kinetic model without magnetohydrodynamical systems nor temperature distributions, fully describing these interactions with a physics informed data-driven structure as a substitution of empirical assumptions.
+That's why, this work departs from that approach, seeking the description of Solar Wind kinetic model without magnetohydrodynamical systems nor temperature distributions, fully describing these interactions with a physics informed data-driven structure as a substitution of empirical assumptions.
 
@@
 
+This is a widely used model to account for anisotropies on Solar Wind \cite{Verscharen_2016} and as a chosen alternative to MHD models \cite{grandin2023hybridvlasovsimulationsoftxray}.
+
 \section{Physics Informed Machine Learning}
 Usually, solely data-driven methods require huge volumes of high-quality data with black box algorithms that prevent interpretability, a key factor for physical modeling.
-The most recent research \cite{PINNS} advocate for physical informed virtual loss function terms, leading off to the vastly known field of Phyics Informed Neural Networks.
+The most recent research advocate for physical informed virtual loss function terms \cite{}, leading off to the vastly known field of Physics Informed Neural Networks.
 
@@
 The most recent research, using PINNs, develop a way to not just learn functions, but to learn functional operators: Lagrangian, Hamiltonian, etc.
-This research \cite{} creates a neural network that takes the general coordinates $q$ and $\dot q$ as input, to provide the lagrangian as an output, and leverage the auto-differentiation capabilities of modern machine learning frameworks to enforce the euler-lagrange equation:
+\cite{cranmer2020lagrangianneuralnetworks, greydanus2019hamiltonianneuralnetworks} The first research creates a neural network that takes the general coordinates $q$ and $\dot q$ as input, to provide the lagrangian as an output, taking advantage of the auto-differentiation capabilities of modern machine learning frameworks to enforce the euler-lagrange equation with gradient-descent methods:
 
@@
 
-As usage of $q$ and $\dot q$ as inputs is very convenient for the auto-differentiation engines, making it further computationally feasible.
+\subsection{Neural Operators for discretization-invariant inference}
 

### 2024-11-12 download_data.py

@@ -1,0 +1,12 @@
+from starstream import DataDownloading, WIND, DSCOVR
+from utils import create_dates
+
+if __name__ == "__main__":
+    DataDownloading(
+        [
+            WIND.TDP_PM(root="/data/Vlasov"),
+            # DSCOVR.FaradayCup(root = '/data/Vlasov'),
+            # DSCOVR.Magnetometer(root =  '/data/Vlasov'),
+        ],
+        create_dates(),
+    )

### 2024-11-12 htuning.py

@@ -5,12 +5,24 @@ from .loss import criterion
 from .model import Model
 import optuna
 
-labels: List[str] | str = criterion().labels
+labels: List[str] | str = criterion().labels  #
 
 
-def objective(trial: optuna.trial.Trial) -> Dict[str, float | int | str]:
-    # Define the objective
-    return hparams
+def objective(trial: optuna.trial.Trial) -> Dict:
+    num_layers: int = trial.suggest_int("num_layers", 1, 5)
+    return dict(
+        optimizer="adam",
+        scheduler="one-cycle",
+        layers=[
+            trial.suggest_int(f"layer_{i}", 1, 128) for i in range(1, num_layers + 1)
+        ],
+        activations=[
+            trial.suggest_categorical(f"activation_{i}", ["ReLU", "Softmax"])
+            for i in range(1, num_layers + 1)
+        ],
+        lr=trial.suggest_float("lr", 1e-5, 5e-2, log=True),
+        weight_decay=trial.suggest_float("weight_decay", 1e-6, 1e-2, log=True),
+    )
 
 
 if __name__ == "__main__":
@@ -22,7 +34,7 @@ if __name__ == "__main__":
         datamodule_kwargs=dict(pin_memory=True, num_workers=8, batch_size=32),
         directions=["minimize" for _ in range(len(labels))],
         precision="high",
-        n_trials=150,
+        n_trials=10000,
         trainer_kwargs=dict(
             logger=True,
             enable_checkpointing=False,
@@ -30,7 +42,7 @@ if __name__ == "__main__":
             accelerator="cuda",
             devices=1,
             log_every_n_steps=22,
-            precision="bf16-mixed",
+            precision="32",
             limit_train_batches=1 / 3,
             limit_val_batches=1 / 3,
         ),

### 2024-11-12 loss.py

@@ -1,15 +1,68 @@
 from lightorch.nn.criterions import LighTorchLoss
 from torch import Tensor
+from typing import Tuple
+from scipy.constants import e, proton_mass
+import torch
 
 
-class criterion(LighTorchLoss):
+class StatisticalMechanicsInformedLoss(LighTorchLoss):
     def __init__(
         self,
+        alpha_1: float,
+        alpha_2: float,
     ) -> None:
         super().__init__(
-            labels=[],
-            factors=[],
+            labels=[
+                "PDE",
+                "CurrentDensityCondition",
+                "LiouvilleCondition",
+                "ProbabilisticCondition",
+                "Entropy",
+            ],
+            factors={
+                "PDE": alpha_1,
+                "BoundaryConditions": alpha_2,
+            },
         )
 
-    def forward(self, x: Tensor) -> Tensor:
-        return super().forward(x)
+    def forward(self, **kwargs) -> Tuple[Tensor, ...]:
+        f_alpha_r = torch.autograd.grad(
+            kwargs["f_alpha"], kwargs["r"], create_graph=True
+        )[0]
+        f_alpha_v = torch.autograd.grad(
+            kwargs["f_alpha"], kwargs["v"], create_graph=True
+        )[0]
+        f_alpha_t = torch.autograd.grad(
+            kwargs["f_alpha"], kwargs["t"], create_graph=True
+        )[0]
+        v_cross_B = torch.cross(kwargs["v"], kwargs["B"], dim=-1)
+
+        v_dot_f_alpha_r = torch.einsum("bsi,bsi->bs", kwargs["v"], f_alpha_r)
+
+        F_m_dot_f_alpha_v = torch.einsum(
+            "bsi,bsi->bs", (e / proton_mass) * (kwargs["E"] + v_cross_B), f_alpha_v
+        )
+
+        vlasov = (f_alpha_t + v_dot_f_alpha_r + F_m_dot_f_alpha_v) ** 2
+
+        integral_rv = torch.sum(kwargs["f_alpha"]) * self.volume_phase
+        j_density = torch.sum(kwargs["f_alpha"] * kwargs["v"]) * self.volume_phase
+
+        current_density = e * j_density
+        liouville = (
+            torch.autograd.grad(integral_rv, kwargs["t"], create_graph=True)[0] ** 2
+        )
+
+        non_reduced_entropy = -kwargs["f_alpha"] * torch.log(kwargs["f_alpha"])
+
+        entropy_criteria = -torch.autograd.grad(
[…]

### 2024-11-12 model.py

@@ -1,23 +1,41 @@
 from lightorch.training.supervised import Module
+from lightorch.nn import DeepNeuralNetwork
 from torch import nn, Tensor
-from .loss import criterion
+from .loss import StatisticalMechanicsInformedLoss
+from typing import Sequence, Union
 
 
-class model_name(nn.Module):
+class PINN(DeepNeuralNetwork):
     def __init__(
         self,
-    ) -> None:
-        super().__init__()
+        in_features: int,
+        layers: Sequence[int],
+        activations: Sequence[Union[str, None]],
+    ):
+        super().__init__(
+            in_features,
+            layers,
+            list(
+                map(
+                    lambda activation: (
+                        getattr(nn, activation) if activation is not None else None
+                    ),
+                    activations,
+                )
+            ),
+        )
 
-    def forward(self, x: Tensor) -> Tensor:
-        return out
+    def forward(self, F: Tensor) -> Tensor:
+        return super().forward(F)
 
 
 class Model(Module):
     def __init__(self, **hparams) -> None:
         super().__init__(**hparams)
-        self.model = model_name()
-        self.criterion = criterion()
+        self.model = PINN(3, hparams["layers"], hparams["activations"])
+        self.criterion = StatisticalMechanicsInformedLoss(
+            hparams["alpha_1"], hparams["alpha_2"]
+        )
 
     def forward(self, x: Tensor) -> Tensor:
         return self.model(x)

### 2024-11-12 utils.py

@@ -1,0 +1,140 @@
+from dateutil.relativedelta import relativedelta
+from typing import List, Union, Tuple, Callable
+from datetime import datetime, timedelta
+from torch import Tensor
+import pandas as pd
+import torch
+import os
+
+
+def get_data_paths() -> List[Callable[[str], str]]:
+    return [
+        lambda date: f"./data/SOHO/CELIAS_PM/{date}.csv",
+        lambda date: f"./data/SOHO/CELIAS_SEM/{date}.csv",
+        lambda date: f"./data/WIND/MAG/{date}.csv",
+        lambda date: f"./data/WIND/TDP_PM/{date}.csv",
+        lambda date: f"./data/ACE/SWEPAM/{date}.csv",
+        lambda date: f"./data/ACE/EPAM/{date}.csv",
+        lambda date: f"./data/ACE/SIS/{date}.csv",
+        lambda date: f"./data/ACE/MAG/{date}.csv",
+        lambda date: f"./data/DSCOVR/L2/faraday/{date}.csv",
+        lambda date: f"./data/DSCOVR/L2/magnetometer/{date}.csv",
+    ]
+
+
+def get_data(paths: List[str]) -> Tensor:
+    df_list: List[pd.DataFrame] = []
+    for path in paths:
+        df_list.append(pd.read_csv(path))
+    return torch.from_numpy(pd.concat(df_list, axis=1).values).unsqueeze(0)
+
+
+def timedelta_to_freq(timedelta_obj) -> str:
+    total_seconds = timedelta_obj.total_seconds()
+
+    if total_seconds % 1 != 0:
+        raise ValueError("Timedelta must represent a whole number of seconds")
+
+    days = total_seconds // (24 * 3600)
+    hours = (total_seconds % (24 * 3600)) // 3600
+    minutes = ((total_seconds % (24 * 3600)) % 3600) // 60
+    seconds = ((total_seconds % (24 * 3600)) % 3600) % 60
+
+    freq_str = ""
+
+    if days > 0:
+        freq_str += str(int(days)) + "day"
+    if hours > 0:
+        freq_str += str(int(hours)) + "hour"
+    if minutes > 0:
+        freq_str += str(int(minutes)) + "min"
+    if seconds > 0:
+        freq_str += str(int(seconds)) + "sec"
+
+    return freq_str
+
+
+def datetime_interval(
+    init: datetime,
+    last: datetime,
+    step_size: Union[relativedelta, timedelta],
[…]

### 2024-11-12 paper/main.tex
paper default

@@
 
-\title{Determining Solar Wind's distribution with Vlasov PINNs}
+\title{Statistical Mechanics Informed Neural Networks for Solar Wind modeling}
 \author{Jorge D. Enciso}
@@
 
-\section{Kinetic Models}
+\section{Kinetic models and general constraints}
 In order to fully assume the magnetohydrodynamical model, interactions must converge to the Maxwell-Boltzmann thermodynamical distribution for certain conditions, implying a high colisionality degree, contrary to empirical observations of the Solar Wind.
@@
 This is a widely used model to account for anisotropies on Solar Wind \cite{Verscharen_2016} and as a chosen alternative to MHD models \cite{grandin2023hybridvlasovsimulationsoftxray}.
+
+\subsection{Liouville's theorem}
+
+\subsection{Entropy}
 

### 2024-11-12 paper/main.tex
paper default
paper

@@
 
-Traditional models and methodologies fail to account for the anysotropic nature of the Solar Wind.
+Traditional models and methodologies neglect the anysotropic nature of the Solar Wind.
 The Magnetohydrodynamics Model assumes a Boltzmann-Maxwellian distribution, doted with standard thermodynamical behavior.
@@
 On the matter of Solar Wind modeling, several numerical approaches are developed to understand near-Sun physics given certain boundary conditions.
-This research \cite{windmodelling1} creates a magnetohydrodynamical modeling scheme to understand these interactions under certain restrictions that approach heliospheric conditions.
-In this case, the current work is seeking more general descriptors of the Solar Wind, with real L1-Lagrange data as a cornerstone for the modeling scheme.
+This research \cite{windmodelling1} creates a magnetohydrodynamical modeling scheme to understand these interactions under certain restrictions that approach the conditions from the transition from the photosphere to the heliospheric distance of 27 solar radii.
+In this case, the current work is seeking more general descriptors of the Solar Wind, with real L1-Lagrange data as a cornerstone for the modeling scheme to finally get a kinetic description of the phenomena.
 
@@
 
-\cite{kumar2023physicsinformedneuralnetworks}
+Regarding Vlasov-Maxwell modeling with physics informed neural networks, this paper constitutes the first machine learning approach intersecting both this theory and the novel PDE solving scheme proposed by PINNs.
+\cite{kumar2023physicsinformedneuralnetworks} It employs the model onto fusion modeling within the colisionless constraint.
+It constitutes an important basis for the present paper, offering a perspective of the efficiency and satisfaction offered by using these methods on alternative fields.
+
+They use the VOICE code to numerically solve the Vlasov-Poisson system of equations, and use this data as ground truth for the Physics Informed Neural Network.
+Given that it's an Integro-Differential equation, they advocate for the usage of I-PINN (integrable-PINN) \cite{} which uses the fundamental theorem of Calculus to approach the integrable part.
+This integration approach will be used by the present work as well to compute several constraints tied to statistical mechanics.
 
@@
 
-\cite{johnson2022physics}
+This research \cite{johnson2022physics} is regarded as one of the first physics informed approaches towards geoeffective solar wind prediction.
+It uses L1-Lagrange spacecraft readings and enforces the Ohm's Law for an ideal plasma, following the Ideal MHD model for space plasma.
+It uses a wide variety of model architectures (GRU, LSTM, 2D CNN ResNets, etc.) to test their effectiveness under this task.
 
@@
 
-\subsection{Boltzmann's equation}
-Within this frame, different kinetic statistical descriptors are suitable for the task as long as they make the correct assumptions.
-Therefore, the usage of Boltzmann's equation works perfectly within certain boundaries:
+\subsection{Boltzmann's Equation}
+
+In the context of kinetic plasma theory, various statistical descriptors are applied, provided they make appropriate assumptions about particle interactions.
+Boltzmann's equation is often used to describe the evolution of the particle distribution function, $f_\alpha(\mathbf{r}, \mathbf{v}, t)$, by considering the effects of particle collisions:
 
 \begin{equation}
-    \frac{\partial f}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha= \left(\frac{\delta f}{\delta t}\right)_{col}
+    \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha = \left(\frac{\delta f}{\delta t}\right)_{\mathrm{col}}
 \end{equation}
 
-This result is widely known for describing the Maxwell-Boltzmann distribution under defined conditions, so an alternative interpretation that accounts for colisionless plasma should be integrated:
+The left side of the equation describes the rate of change of $f_\alpha$ due to the particles' motion and external forces, while the term $\left( \frac{\delta f_\alpha}{\delta t} \right)_{col}$on the right side represents the change in $f_\alpha$ due to collisions.
 
-\subsection{Vlasov's equation}
-The Vlasov's equation is the colisionless variant of the Boltzmann's equation. Therefore, it is a  suitable kinetic model for the Solar Wind, accounting for key properties of it.
+This formulation is essential for systems where collisions are frequent and lead to a Maxwell-Boltzmann distribution under equilibrium conditions.
+However, for plasmas like the solar wind, where collisions are rare, a collisionless description is more appropriate.
+
+\subsection{Vlasov's Equation}
+
+The Vlasov equation is the collisionless analog of the Boltzmann equation, making it a more suitable kinetic model for the solar wind.
+It describes the evolution of the particle distribution function in a collisionless plasma, where interactions are primarily governed by external forces rather than particle collisions:
+
+\begin{equation} \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \frac{\mathbf{F}}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0 \end{equation}
+
+In this context, the system is typically dominated by electromagnetic interactions. Thus, substituting with the Lorentz force is the ideal case:
 
 \begin{equation}
-    \frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{\mathbf{F}}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
+    \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \frac{q}{m} (\mathbf{E} + \mathbf{v} \times \mathbf{B}) \cdot \nabla_{\mathbf{v}} f_\alpha = 0
 \end{equation}
 
-In this case, governed by electromagnetic interactions. Therefore, replacing the force $\mathbf{F}$ with the Lorentz Force:
+This form of Vlasov equation is widely used to capture the dynamics of collisionless plasmas, as it allows us to account for anisotropies in the solar wind, such as those arising from velocity space instabilities and non-thermal distributions of particle energies \cite{Verscharen_2016}.
+Unlike ideal MHD models, which assume that the Lorentz force term can be neglected, the Vlasov equation includes this term to account for deviations from isotropy and Maxwellian distributions in the solar wind, requiring numerical approximation techniques for practical solutions \cite{grandin2023hybridvlasovsimulationsoftxray}.
+
+Note that Vlasov's equation is tightly related to statistical mechanics through Liouville's theorem:
 
 \begin{equation}
-    \frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
+    \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = 0
 \end{equation}
 
-This is a widely used model to account for anisotropies on Solar Wind \cite{Verscharen_2016} and as a chosen alternative to MHD models \cite{grandin2023hybridvlasovsimulationsoftxray}.
+If we compute use the chain rule along the phase and time space, we end up with the following expression:
+\begin{equation}
+    \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = \frac{\partial f_\alpha}{\partial t} + \nabla_{\mathbf{r}} f_\alpha \frac{\partial \mathbf{r}}{\partial t} + \nabla_{\mathbf{v}} f_\alpha \frac{\partial \mathbf{v}}{\partial t} = 0
+\end{equation}
 
-\subsection{Liouville's theorem}
+which indeed is:
+
+\begin{equation}
+    \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
+\end{equation}
 
 \subsection{Entropy}
+
+According to the second law of thermodynamics, the entropy of a closed system cannot decrease over time; it may increase or remain constant.
+In the kinetic framework, the entropy production rate is given by the time derivative of the entropy functional:
+
+\begin{equation}
+    0 \leq \frac{\mathrm{d}}{\mathrm{dt}} \int f_\alpha \log f_\alpha d^3\mathbf{r} d^3\mathbf{v}
+\end{equation}
+
+where $\f_\alpha \log f_\alpha$ represents the local entropy density in phase space.
+In a collisionless plasma, this entropy can remain constant if the system maintains phase-space coherence, but under realistic conditions, such as wave-particle interactions, entropy may increase as energy cascades through different scales.
+This formalism highlights how irreversible processes in the plasma lead to an increase in degrees of freedom, a key consideration for accurately modeling solar wind behavior.
+
+For this case particularly, the colisionless nature of the plasma and Liouville's theorem assumes no change in entropy, if we use Leibniz method and compute the time derivative of the integrant:
+
+\begin{equation}
+    \frac{d}{dt} f_\alpha \log f_\alpha = \frac{df_\alpha}{dt} \log f_\alpha + \frac{f_\alpha}{f_\alpha}\frac{df_\alpha}{dt} = 0 + 0 = 0
+\end{equation}
 
@@
 
+The current work extrapolates this concept to the field of density functions, embedding it into the training phase by ensuring the normalized nature of a density function, the second law of thermodynamics, and systems governed by kinetic theory following Liouville's theorem.
+
 \subsection{Neural Operators for discretization-invariant inference}
+
+Recent research on operator learning lead to the creation of Physics informed Neural Operators (PINOs), a set of weighted integral operators that are sequencially evaluated through the network.
 

### 2024-11-13 paper/main.tex
default

--- 2024-11-12 paper/main.tex
@@
 
-\cite{kumar2023physicsinformedneuralnetworks}
+Regarding Vlasov-Maxwell modeling with physics informed neural networks, this paper constitutes the first machine learning approach intersecting both this theory and the novel PDE solving scheme proposed by PINNs.
+\cite{kumar2023physicsinformedneuralnetworks} It employs the model onto fusion modeling within the colisionless constraint.
 

### 2024-11-13 download_data.py
checked blacked

@@ -4,9 +4,9 @@ from utils import create_dates
 if __name__ == "__main__":
     DataDownloading(
         [
-            WIND.TDP_PM(root="/data/Vlasov"),
-            # DSCOVR.FaradayCup(root = '/data/Vlasov'),
-            # DSCOVR.Magnetometer(root =  '/data/Vlasov'),
+            WIND.TDP_PM(root="/data/Vlasov/TDP_PM"),
+            DSCOVR.FaradayCup(root="/data/Vlasov/DSCOVR"),
+            DSCOVR.Magnetometer(root="/data/Vlasov/DSCOVR"),
         ],
         create_dates(),
     )

### 2024-11-13 paper/main.tex
default

--- 2024-11-12 paper/main.tex
@@
 
-Several attempts to model Solar Wind's dynamics can be empirically pinpointed : analytical methods \cite{BLUME202396}, numerical modelling \cite{10.3389/fspas.2023.1105797, windmodelling1, Gombosi_2018}, and machine learning approaches \cite{comp_2, comp_3, guastavino2024forecastinggeoffectiveeventssolar, sabbatini2023solarwindspeedestimate, https://doi.org/10.1029/2023SW003561}.
+Several attempts to model Solar Wind's dynamics can be empirically pinpointed: analytical methods \cite{BLUME202396}, numerical modelling \cite{10.3389/fspas.2023.1105797, windmodelling1, Gombosi_2018}, and machine learning approaches \cite{comp_2, comp_3, guastavino2024forecastinggeoffectiveeventssolar, sabbatini2023solarwindspeedestimate, https://doi.org/10.1029/2023SW003561}.
 All of them with different purposes.
 
-Traditional models and methodologies neglect the anysotropic nature of the Solar Wind.
-The Magnetohydrodynamics Model assumes a Boltzmann-Maxwellian distribution, doted with standard thermodynamical behavior.
-On the other hand, the Solar Wind is a collisionless plasma, precluding its inclusion solely within those boundaries.
-Thus, the present work emphasizes the usage of learnable neural operators engineered to analytically and empirically fit kinetic descriptions.
+The solar wind presents unique challenges due to its deviation from classical assumptions.
+It is a quasy-neutral, colisionless, and empirically quasy-isotropic ionized plasma.
+\cite{Verscharen2022} Therefore, the adaptation of plasma dynamics frameworks for solar wind modeling must be carefully analyzed.
+For instance, the ideal magnetohydrodynamics (Ideal MHD) model is used under the assumption of an ionized plasma that converges to the Boltzmann distribution (Maxwell-Boltzmann distribution for temperatures), but it's an empirically unfeasible distribution as it misrepresents historical observations on solar wind's behavior.
 
-Moreover, kinetic frames based on Maxwellian electromagnetism tend to be computationally expensive for the microscopic and macroscopic frame:
+On the other hand, using numerical approximators to model plasma dynamics ends up being computationally expensive. Take the Ideal MHD as an example:
 
 \begin{equation}
-    n_{\alpha} (\mathbf{r}, t) := \int f_{\alpha}(\mathbf{r}, \mathbf{v}, t) d^3v
+    \nabla \cdot \textbf{E} = \frac{\sigma}{\epsilon_0}
+\end{equation}
+\begin{equation}
+    \nabla \cdot \textbf{B} = 0
+\end{equation}
+\begin{equation}
+    \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0
+\end{equation}
+\begin{equation}
+    \frac{d}{dt} \left( \frac{p}{\rho^\gamma} \right)= 0
+\end{equation}
+\begin{equation}
+    \textbf{E} + \textbf{v} \times \textbf{B} = 0
+\end{equation}
+\begin{equation}
+    \frac{\partial \textbf{B}}{\partial t} = \nabla \times (\textbf{v} \times \textbf{B})
+\end{equation}
+\begin{equation}
+    \textbf{J} \times \textbf{B} = \frac{(\textbf{B}\cdot \nabla) \textbf{B}}{\mu_0} - \nabla \left(\frac{B^2}{2\mu_0}\right)
 \end{equation}
 
-\begin{equation}
-    \tau(r, t) := \sum q_{\alpha} n_{\alpha}
-\end{equation}
+where $\mathbf{E}$ is the electric field, $\sigma$ the charge density, $\mathbf{B}$ the magnetic field, $p$ the pressure, $\rho$ the mass density, $v$ the general velocity, and $\gamma $ the adiabatic index.
 
-\begin{equation}
-    \mathbf{u_{\alpha}} (\mathbf{r}, t) := \frac{1}{n_{\alpha}(\mathbf{r}, t)} \int v f_{\alpha}(\mathbf{r}, v, t) d^3v
-\end{equation}
+Nevertheless, recent research on machine learning computational methods enable novel approaches towards PDE solving and operator learning.
+These architectures, named Physics Informed Neural Networks, allow learnable solutions for less computational price thanks to modern hardware optimizations and shallower representations with universal approximators (in this case, neural networks).
+Hence, the scope of this research is to adhere these concepts on a deeper basis, proposing a new methodology within numerical statistical mechanics for solar wind modeling.
 
-\begin{equation}
-    j(\mathbf{r}, t) := \sum q_{\alpha} n_{\alpha} \mathbf{u_{\alpha}}
-\end{equation}
-
-where $n_{\alpha}$ is the particle density, $u_{\alpha}$ the particle's drift velocity, $\tau(r, t)$ is the charge density, $j(r, t)$ is the current density, and $f_{\alpha}(r, v, t)$ is the Boltzmann's density function.
-
-Yet modern physics informed machine learning research schemes enable faster inference time and less computational load: embedding systems' equations within the loss function and constraining network's backbone based on empirical approaches.
+To this date, the literature doesn't embark on statistical mechanics informed neural networks for probability density function learning on a solar wind scheme.
+Therefore, the objective of this work is to leverage statistical mechanics theory and physics informed neural networks to propose a new numerical approximation for the kinetic behavior of the solar wind.
 
@@
 It uses a wide variety of model architectures (GRU, LSTM, 2D CNN ResNets, etc.) to test their effectiveness under this task.
+
+\subsection{Statistical-Physics-Informed Neural Networks (Stat-PINNs): A Machine Learning Strategy for Coarse-graining Dissipative Dynamics}
 
@@
 
-This models where created as numerical methods for solving Partial Differential Equations that embed physical modeling.
+This models were created as numerical methods for solving Partial Differential Equations that embed physical modeling.
 It further extended to all fields that required PDE solving as a computationally efficient numerical alternative to other solvers.
@@
 The current work extrapolates this concept to the field of density functions, embedding it into the training phase by ensuring the normalized nature of a density function, the second law of thermodynamics, and systems governed by kinetic theory following Liouville's theorem.
-
-\subsection{Neural Operators for discretization-invariant inference}
-
-Recent research on operator learning lead to the creation of Physics informed Neural Operators (PINOs), a set of weighted integral operators that are sequencially evaluated through the network.
 
@@
 The neural network is trained using the standard gradient-descent based optimization algorithm: Adam.
-One Cycle learning optimizer is used to facilitate the learning rate plateau.
-Weight decay is implemented as a regularizer.
+One Cycle learning rate scheduler imizer is used to facilitate the convergence to a learning rate plateau.
+L2 regularizer and dropout is implemented to avoid overfitting.
 
-The loss function consists on a virtual term, three integral boundary condition terms, and a l2 regularizer term:
+The loss function consists on a Liouville, probabilistic, and entropy term.
 
 \begin{equation}
-    \mathcal{L}_{virtual}\left(\mathbf{E}, \mathbf{B}, \mathbf{v}, \mathbf{r}; \mathbf{\theta}\right) = \left(\frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha \right)^2
+    \mathcal{L}_{Liouville}\left(\mathbf{E}, \mathbf{B}, \mathbf{v}, \mathbf{r}; \mathbf{\theta}\right) = \left(\frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha \right)^2
 \end{equation}
@@
 \begin{equation}
-    \mathcal{L}_{BC_1} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} d^3\mathbf{r} - 1\right)^2
+    \mathcal{L}_{prob} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} d^3\mathbf{r} - 1\right)^2
 \end{equation}
@@
 \begin{equation}
-    \mathcal{L}_{BC_2} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} - n_\alpha(\mathbf{r}, t) \right)^2
-\end{equation}
-
-\begin{equation}
-    \mathcal{L}_{BC_3} \left(x; \mathbf{\theta}\right) = \left(\sum_{\alpha} q_\alpha n_\alpha  \mathbf{u}_\alpha- j(\mathbf{r}, t) \right)^2
+    \mathcal{L}_S(\mathbf{r}, \mathbf{v}, t) = \left(\frac{d}{dt}\int f_\alpha \log f_\alpha \right)^2
 \end{equation}
@@
 \begin{equation}
-    \mathcal{L}\left[\mathcal{L}_{virtual}, \mathcal{L}_{BC}, \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{virtual} + \alpha_2 \sum_{i} \mathcal{L}_{BC_i} + \lambda \mathcal{L}_{regularizer}
+    \mathcal{L}\left[\mathcal{L}_{Liouville}, \mathcal{L}_{prob}, \mathcal{L}_{S} \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{Liouville} + \alpha_2 \mathcal{L}_{prob} + \alpha_3 \mathcal{L}_{S} + \lambda \mathcal{L}_{regularizer}
 \end{equation}
@@
 \section{Conclusion}
-This work demonstrated the effectiveness of Physics-Informed Neural Networks to model Solar Wind's behavior tied to diverse empirical and analytical formulations.
+This work demonstrated the effectiveness of Physics-Informed Neural Networks to model Solar Wind's behavior tied to diverse statistical mechanics formulations that lead to Vlasov's kinetic model.
 By embedding the governing physical equations directly into the loss function, we can efficiently train a neural network to approximate the solution of this complex, non-linear system.
-Future work will focus on improving the kernel design and extending the approach to more sophisticated MHD scenarios, such as the interactions between Solar Wind and Sun's magnetosphere.
+
+Future work will focus on improving the architecture design with Physics informed Neural Operators (PINOs) \cite{li2023physicsinformedneuraloperatorlearning} by embedding Maxwell's equations for data continuization through differentiable splines, the satellite data dependance inhibits the introduction os such powerful frameworks in this case.
 

### 2024-11-13 paper/main.tex
default paper

--- 2024-11-12 paper/main.tex
@@
 \begin{equation}
-    \mathcal{L}_{virtual}\left(\mathbf{E}, \mathbf{B}, \mathbf{v}, \mathbf{r}; \mathbf{\theta}\right) = \left(\frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha \right)^2
+    \mathcal{L}_{liouville}\left(\mathbf{E}, \mathbf{B}, \mathbf{v}, \mathbf{r}; \mathbf{\theta}\right) = \left(\frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha \right)^2
 \end{equation}
@@
 \begin{equation}
-    \mathcal{L}\left[\mathcal{L}_{virtual}, \mathcal{L}_{BC}, \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{virtual} + \alpha_2 \sum_{i} \mathcal{L}_{BC_i} + \lambda \mathcal{L}_{regularizer}
+    \mathcal{L}\left[\mathcal{L}_{liouville}, \mathcal{L}_{prob}, \mathcal{L}_{S} \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{liouville} + \alpha_2 \mathcal{L}_{prob} + \alpha_3 \mathcal{L}_{S} + \lambda \mathcal{L}_{regularizer}
 \end{equation}

### 2024-11-14 download_data.py
default

@@ -1,12 +1,21 @@
-from starstream import DataDownloading, WIND, DSCOVR
+from starstream import DataDownloading, WIND, DSCOVR, ACE
 from utils import create_dates
 
 if __name__ == "__main__":
-    DataDownloading(
-        [
-            WIND.TDP_PM(root="/data/Vlasov/TDP_PM"),
-            DSCOVR.FaradayCup(root="/data/Vlasov/DSCOVR"),
-            DSCOVR.Magnetometer(root="/data/Vlasov/DSCOVR"),
-        ],
-        create_dates(),
-    )
+    for tuple_date in create_dates():
+        try:
+            DataDownloading(
+                [
+                    WIND.TDP_PM(root="/data/Vlasov/TDP_PM"),
+                    DSCOVR.FaradayCup(root="/data/Vlasov/DSCOVR/FaradayCup"),
+                    DSCOVR.Magnetometer(root="/data/Vlasov/DSCOVR/Magnetometer"),
+                    ACE.SWEPAM(download_path="/data/Vlasov/ACE/SWEPAM"),
+                    ACE.SWICS(download_path="/data/Vlasov/ACE/SWICS"),
+                    ACE.EPAM(download_path="/data/Vlasov/ACE/EPAM"),
+                    ACE.MAG(download_path="/data/Vlasov/ACE/MAG"),
+                    ACE.SIS(download_path="/data/Vlasov/ACE/SIS"),
+                ],
+                tuple_date,
+            )
+        except:
+            continue

### 2024-12-09 paper/jmlr/sample.tex
default

--- 2024-11-13 paper/main.tex
@@
-\documentclass[12pt]{article}
+\documentclass[twoside,11pt]{article}
 
+\usepackage{blindtext}
 \usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{cite}
 
-\title{Statistical Mechanics Informed Neural Networks for Solar Wind modeling}
-\author{Jorge D. Enciso}
-\date{\today}
+% Any additional packages needed should be included after jmlr2e.
+% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
+% and defines many common macros, such as 'proof' and 'example'.
+%
+% It also sets the bibliographystyle to plainnat; for more information on
+% natbib citation styles, see the natbib documentation, a copy of which
+% is archived at http://www.jmlr.org/format/natbib.pdf
+
+% Available options for package jmlr2e are:
+%
+%   - abbrvbib : use abbrvnat for the bibliography style
+%   - nohyperref : do not load the hyperref package
+%   - preprint : remove JMLR specific information from the template,
+%         useful for example for posting to preprint servers.
+%
+% Example of using the package with custom options:
+%
+% \usepackage[abbrvbib, preprint]{jmlr2e}
+
+\usepackage{jmlr2e}
+
+\newcommand{\dataset}{{\cal D}}
+\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
+\newcommand{\diff}[2]{\frac{d #1}{d #2}}
+
+\usepackage{lastpage}
+\jmlrheading{25}{2024}{1-\pageref{LastPage}}{12/24; Revised ?/25}{?/25}{21-0000}{Jorge D. Enciso}
+
+% Short headings should be running head and authors last names
+
+\ShortHeadings{Statistical Mechanics Informed Neural Networks for Solar Wind modeling}{Jorge D. Enciso}
+\firstpageno{1}
 
@@
 
+\title{Statistical Mechanics Informed Neural Networks for Solar Wind modeling}
+
+\author{\name Jorge D. Enciso \email jorged.encyso@gmail.com \\
+    \addr Asuncion, Paraguay}
+
+\editor{My editor}
+
 \maketitle
 
-\begin{abstract}
-    %TODO
+\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
+The solar wind, a collisionless, quasi-neutral plasma, presents unique challenges for modeling due to its deviation from classical assumptions such as the Maxwell-Boltzmann distribution.
+Traditional numerical and analytical approaches are computationally expensive and often rely on oversimplified empirical constraints.
+Recent advancements in machine learning, particularly Physics-Informed Neural Networks (PINNs), offer an efficient alternative for solving partial differential equations while incorporating physical laws.
+This research introduces a novel framework: Statistical Mechanics Informed Neural Networks, tailored for modeling the kinetic behavior of the solar wind.
+By leveraging statistical mechanics and avoiding restrictive magnetohydrodynamic assumptions, the proposed methodology directly approximates the solar wind's probability density function.
+Using data from near-Earth L1-Lagrange point observations, this work demonstrates the feasibility of PINNs in accurately capturing the solar wind's anisotropic and non-thermal dynamics.
+The findings highlight the potential of integrating statistical physics and machine learning for advanced space plasma modeling, paving the way for more precise solar-terrestrial interaction predictions.
 \end{abstract}
 
-\tableofcontents
-\newpage
+\begin{keywords}
+Physics informed Neural Networks, Solar Wind, Statistical Mechanics, Vlasov's Equation, Maxwell Equations, Ideal MHD
+\end{keywords}
 
@@
 
-\begin{equation}
-    \nabla \cdot \textbf{E} = \frac{\sigma}{\epsilon_0}
-\end{equation}
-\begin{equation}
-    \nabla \cdot \textbf{B} = 0
-\end{equation}
-\begin{equation}
-    \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0
-\end{equation}
-\begin{equation}
-    \frac{d}{dt} \left( \frac{p}{\rho^\gamma} \right)= 0
-\end{equation}
-\begin{equation}
-    \textbf{E} + \textbf{v} \times \textbf{B} = 0
-\end{equation}
-\begin{equation}
-    \frac{\partial \textbf{B}}{\partial t} = \nabla \times (\textbf{v} \times \textbf{B})
-\end{equation}
-\begin{equation}
-    \textbf{J} \times \textbf{B} = \frac{(\textbf{B}\cdot \nabla) \textbf{B}}{\mu_0} - \nabla \left(\frac{B^2}{2\mu_0}\right)
-\end{equation}
+\begin{align*}
+    \nabla \cdot \textbf{E} &= \frac{\sigma}{\epsilon_0} \\
+    \nabla \cdot \textbf{B} &= 0 \\
+    \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &= 0 \\
+    \frac{d}{dt} \left( \frac{p}{\rho^\gamma} \right) &= 0 \\
+    \textbf{E} + \textbf{v} \times \textbf{B} &= 0 \\
+    \frac{\partial \textbf{B}}{\partial t} &= \nabla \times (\textbf{v} \times \textbf{B}) \\
+    \textbf{J} \times \textbf{B} &= \frac{(\textbf{B}\cdot \nabla) \textbf{B}}{\mu_0} - \nabla \left(\frac{B^2}{2\mu_0}\right)
+\end{align*}
 
@@
 
-\begin{equation}
+\[
     \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha = \left(\frac{\delta f}{\delta t}\right)_{\mathrm{col}}
-\end{equation}
+\]
 
@@
 
-\begin{equation} \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \frac{\mathbf{F}}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0 \end{equation}
+\[
+    \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \frac{\mathbf{F}}{m} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
+\]
 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:liouville}
     \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \frac{q}{m} (\mathbf{E} + \mathbf{v} \times \mathbf{B}) \cdot \nabla_{\mathbf{v}} f_\alpha = 0
@@
 
-\begin{equation}
+\[
 \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = 0
-\end{equation}
+\]
 
 If we compute use the chain rule along the phase and time space, we end up with the following expression:
-\begin{equation}
+\[
     \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = \frac{\partial f_\alpha}{\partial t} + \nabla_{\mathbf{r}} f_\alpha \frac{\partial \mathbf{r}}{\partial t} + \nabla_{\mathbf{v}} f_\alpha \frac{\partial \mathbf{v}}{\partial t} = 0
-\end{equation}
+\]
 
@@
 
-\begin{equation}
+\[
     \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
-\end{equation}
+\]
 
@@
 
-\begin{equation}
+\[
     0 \leq \frac{\mathrm{d}}{\mathrm{dt}} \int f_\alpha \log f_\alpha d^3\mathbf{r} d^3\mathbf{v}
-\end{equation}
+\]
 
@@
 For this case particularly, the colisionless nature of the plasma and Liouville's theorem assumes no change in entropy, if we use Leibniz method and compute the time derivative of the integrant:
+\[
+    \frac{d}{dt} f_\alpha \log f_\alpha = \frac{df_\alpha}{dt} \log f_\alpha + \frac{f_\alpha}{f_\alpha}\frac{df_\alpha}{dt} = 0 + 0 = 0
+\]
 
-\begin{equation}
-    \frac{d}{dt} f_\alpha \log f_\alpha = \frac{df_\alpha}{dt} \log f_\alpha + \frac{f_\alpha}{f_\alpha}\frac{df_\alpha}{dt} = 0 + 0 = 0
+Thus, the entropy of this system is defined as follows:
+
+\begin{equation} \label{eq:entropy}
+    \frac{\mathrm{d}}{\mathrm{dt}} \int f_\alpha \log f_\alpha d^3\mathbf{r} d^3\mathbf{v} = 0
 \end{equation}
@@
 
-\begin{equation}
+\[
     \frac{d}{dt} \frac{\partial \mathcal{L}}{\partial \dot q} - \frac{\partial \mathcal{L}}{\partial q} = 0
-\end{equation}
+\]
 
@@
 
-The loss function consists on a Liouville, probabilistic, and entropy term.
+The loss function consists on Liouville's theorem (Equation \ref{eq:liouville}), the entropy constraint (Equation \ref{eq:entropy}), a normalization term for probability density functions, and a regularizer term.
 
@@
 \begin{equation}
-    \mathcal{L}_{prob} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} d^3\mathbf{r} - 1\right)^2
+    \mathcal{L}_{P} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} d^3\mathbf{r} - 1\right)^2
 \end{equation}
@@
 \begin{equation}
-    \mathcal{L}\left[\mathcal{L}_{Liouville}, \mathcal{L}_{prob}, \mathcal{L}_{S} \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{Liouville} + \alpha_2 \mathcal{L}_{prob} + \alpha_3 \mathcal{L}_{S} + \lambda \mathcal{L}_{regularizer}
+    \mathcal{L}\left[\mathcal{L}_{Liouville}, \mathcal{L}_{P}, \mathcal{L}_{S}, \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{Liouville} + \alpha_2 \mathcal{L}_{P} + \alpha_3 \mathcal{L}_{S} + \lambda \mathcal{L}_{regularizer}
 \end{equation}
@@
 
-\bibliographystyle{plain}
+\acks{}
 
-\bibliography{references}
+\newpage
+
+\appendix
+
+\section{}
+
+\vskip 0.2in
+\bibliography{}
+
 \end{document}

### 2024-12-09 paper/main.tex
hola

--- 2024-11-13 paper/main.tex
@@
 
-\subsection{Statistical-Physics-Informed Neural Networks (Stat-PINNs): A Machine Learning Strategy for Coarse-graining Dissipative Dynamics}
-
 \section{Kinetic models and general constraints}
@@
 
-\subsection{Boltzmann's Equation}
+\subsection{Boltzmann-Vlasov Equations}
 
@@
 However, for plasmas like the solar wind, where collisions are rare, a collisionless description is more appropriate.
-
-\subsection{Vlasov's Equation}
 
@@
 
-For this case particularly, the colisionless nature of the plasma and Liouville's theorem assumes no change in entropy, if we use Leibniz method and compute the time derivative of the integrant:
+For this case particularly, the collisionless nature of the plasma and Liouville's theorem assumes no change in entropy, if we use Leibniz method and compute the time derivative of the integrant:
 
@@
 
+\subsection{Maxwell's equations}
+
+To complete the description of the solar wind, the Maxwell's electrodynamic formulation provides our final description to well-round the kinetic analysis of the plasma:
+
+\begin{align*}
+    \nabla \cdot E &= \frac{\sum_{\alpha} \rho_\alpha}{\epsilon_0} \\
+    \nabla \cdot B &= 0 \\
+    \nabla \times E &= - \frac{\partial B}{\partial t} \\
+    \nabla \times B &= \mu_0 \left(J + \epsilon_0 \frac{\partial E}{\partial t}\right)
+\end{align*}
+
+\subsection{Complete dynamical system}
+For some general universal approximator $f_n$, also interpreted as dynamical descriptor in this context, there is a parameter space $\theta_n$ tied to the governing constraints ($g_i$) for its respective phenomena:
+
+\begin{align*}
+    f_n(\mathbf{r}, \mathbf{v}, t; \theta_n): \nabla f_n = \sum_{i} \lambda_i \nabla g_i
+\end{align*}
+
+For this sake, we employ a Statistical mechanics informed Vlasov-Maxwell model that fully describes the known, ideal behavior of the solar wind:
+
+\begin{align*}
+    \frac{\partial f_\alpha}{\partial t} + f_v \cdot \nabla_{\mathbf{r}} f_\alpha &+ \frac{q_\alpha ( f_E + f_v \times f_B)}{m_\alpha} \cdot \nabla_{\mathbf{v}} f_\alpha = 0 \text{ (Vlasov)}\\
+    \sum_{\alpha} q_\alpha \int &f_\alpha d^3(f_v) = 0 \text{ (Quasineutrality)}\\
+\frac{d}{dt} \int &f_\alpha \log f_\alpha d^3r d^3(f_v) = 0 \text{ (Liouville)}\\
+\nabla \cdot f_E &= \frac{\sum_{\alpha} \rho_\alpha}{\epsilon_0} \\
+\nabla \cdot f_B &= 0 \\
+\nabla \times f_E &= - \frac{\partial f_B}{\partial t} \\
+\nabla \times f_B &= \mu_0 \left(J + \epsilon_0 \frac{\partial f_E}{\partial t}\right) \text{ (Maxwell)} \\
+\end{align*}
+
+$f_\alpha$ is the kinetic description of each specie, $f_E$ is the electric field descriptor, $f_B$ is the magnetic field descriptor, $f_v$ is the velocity descriptor, $J$ is the current density (derived from $f_\alpha$), and $\rho_\alpha$ is the charge density (derived from $f_\alpha$).
+
+The cross-dependance between both models for the backpropagation process is crucial; thus, a backbone architecture for the electromagnetic descriptor engineered to exploit the backpropagation of the kinetic model can lead to faster convergence in addition to the Maxwell constraints.
+
+A cross-training process that relies on the bonds between the Maxwell and Vlasov PDE criterions will be critical; the PDE loss APORTE to the training process will be gradually incremented through epochs, and inversely correlated with the distances from the boundary contour delimiting the training space to minimize the error margin of non-representative results in the middle of the training process.
+
+This structure provides an optimization process overtly relying on PDE residuals; therefore, an appropiate set of boundary conditions is crucial for the convergence of the physics informed neural networks.
+
+\subsubsection{Scope and boundary conditions}
+The current work utilizes data from L1-Lagrange readings; accordingly, the phase space is constrained to fit predictions up to the Earth's position. As so, we define our parameter space:
+
+\begin{align*}
+    r &\in \left[10R_E, 240R_E\right] \\
+    v_x &\in \left[300, 1000\right]\frac{km}{s} \\
+    t &\in \left[0, 7200]s \\
+\end{align*}
+
+For the kinetic and electormagnetic model, we define the following Dirichlet boundary conditions:
+
+\begin{align*}
+    f_\alpha\left(r, v_{max}, t) &= 0 \\
+    f_\alpha\left(r, v_{min}, t) &= 0 \\
+    f_v\left(r_{min}, t) &= \mathbf{v}_{L1}(t) \\
+    f_v\left(r_{max}, t) &= \mathbf{v}_{Earth}(t) \\
+    f_E\left(r_{min}, t\right) &= E_{L1}(t) \\
+    f_E\left(r_{max}, t\right) &= E_{Earth}(t) \\
+    f_B\left(r_{min}, t\right) &= B_{L1}(t) \\
+    f_B\left(r_{max}, t\right) &= B_{Earth}(t)
+\end{align*}
+
 \section{Physics Informed Machine Learning}
@@
 
-This models were created as numerical methods for solving Partial Differential Equations that embed physical modeling.
+These models were created as numerical methods for solving Partial Differential Equations that embed physical modeling.
 It further extended to all fields that required PDE solving as a computationally efficient numerical alternative to other solvers.
@@
 The current work extrapolates this concept to the field of density functions, embedding it into the training phase by ensuring the normalized nature of a density function, the second law of thermodynamics, and systems governed by kinetic theory following Liouville's theorem.
+
+This model establishes a cornerstone for solar wind modeling, bonding the kinetic model distribution to the results of the electromagnetic model, extending the training space from     L1 Lagrange to a whole grid interval from L1 to the Earth.
 
@@
 
-\subsection{Loss function}
-The neural network is trained using the standard gradient-descent based optimization algorithm: Adam.
-One Cycle learning rate scheduler imizer is used to facilitate the convergence to a learning rate plateau.
-L2 regularizer and dropout is implemented to avoid overfitting.
-
-The loss function consists on a Liouville, probabilistic, and entropy term.
-
-\begin{equation}
-    \mathcal{L}_{Liouville}\left(\mathbf{E}, \mathbf{B}, \mathbf{v}, \mathbf{r}; \mathbf{\theta}\right) = \left(\frac{\partial f_{\alpha}}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}}  f_{\alpha} + \frac{q (\mathbf{E} + \mathbf{v} \times \mathbf{B})}{m} \cdot \nabla_{\mathbf{v}} f_\alpha \right)^2
-\end{equation}
-
-\begin{equation}
-    \mathcal{L}_{prob} \left(x; \mathbf{\theta}\right) = \left(\int_{x \sim X} f_\alpha (\mathbf{v}, \mathbf{r}, t) d^3\mathbf{v} d^3\mathbf{r} - 1\right)^2
-\end{equation}
-
-\begin{equation}
-    \mathcal{L}_S(\mathbf{r}, \mathbf{v}, t) = \left(\frac{d}{dt}\int f_\alpha \log f_\alpha \right)^2
-\end{equation}
-
-\begin{equation}
-    \mathcal{L}_{regularizer} \left( \mathbf{\theta} \right) = \mathbb{E}_{w \in \mathbf{\theta}}\left[ w ^ 2 \right]
-\end{equation}
-
-Getting our final objective function:
-
-\begin{equation}
-    \mathcal{L}\left[\mathcal{L}_{Liouville}, \mathcal{L}_{prob}, \mathcal{L}_{S} \mathcal{L}_{regularizer}\right] = \alpha_1 \mathcal{L}_{Liouville} + \alpha_2 \mathcal{L}_{prob} + \alpha_3 \mathcal{L}_{S} + \lambda \mathcal{L}_{regularizer}
-\end{equation}
-
 \section{Results}

### 2024-12-09 loss.py

@@ -1,68 +1,90 @@
 from lightorch.nn.criterions import LighTorchLoss
 from torch import Tensor
-from typing import Tuple
+from typing import Callable, Tuple
 from scipy.constants import e, proton_mass
 import torch
+from torchquad import MonteCarlo, set_up_backend
 
+set_up_backend('torch', data_type = 'float32')
 
 class StatisticalMechanicsInformedLoss(LighTorchLoss):
     def __init__(
         self,
         alpha_1: float,
         alpha_2: float,
+        alpha_3: float,
+        alpha_4: float,
     ) -> None:
         super().__init__(
             labels=[
-                "PDE",
-                "CurrentDensityCondition",
-                "LiouvilleCondition",
-                "ProbabilisticCondition",
-                "Entropy",
+                "Vlasov",
+                "Liouville",
+                "Quasyneutrality",
+                "Maxwell",
             ],
             factors={
-                "PDE": alpha_1,
-                "BoundaryConditions": alpha_2,
+                "Vlasov": alpha_1,
+                "Liouville": alpha_2,
+                "Quasyneutrality": alpha_3,
+                "Maxwell": alpha_4,
             },
         )
+        self.integral = MonteCarlo()
+
+    def vlasov(self, f: Tensor, r: Tensor, v: Tensor, t: Tensor, E: Tensor, B: Tensor) -> Tensor:
+        """
+        This is the solution for the alpha specie. The solution to the overall
+        solar wind kinetic model is given by the solution to each specie separately.
+        """
+        f_r = torch.autograd.grad(
+            f, r, create_graph=True
+        )[0]
+        f_v = torch.autograd.grad(
+            f, v, create_graph=True
+        )[0]
+        f_t = torch.autograd.grad(
+            f, t, create_graph=True
+        )[0]
+        v_cross_B = torch.cross(v, B, dim=-1)
+
+        v_dot_f_r = torch.einsum("bsi,bsi->bs", v, f_r)
+
+        lorentz_m_dot_f_v = torch.einsum(
+            "bsi,bsi->bs", (e / proton_mass) * (E + v_cross_B), f_v
+        )
+
+        return (f_t + v_dot_f_r + lorentz_m_dot_f_v) ** 2
+
+    def entropy(self, model: Callable) -> Tensor:
+        return self.integral.integrate(
[…]

### 2024-12-09 loss.py
default

@@ -1,7 +1,7 @@
 from lightorch.nn.criterions import LighTorchLoss
 from torch import Tensor
 from typing import Callable, Tuple
-from scipy.constants import e, proton_mass
+from scipy.constants import e, m_e, m_p, mu_0, epsilon_0
 import torch
 from torchquad import MonteCarlo, set_up_backend
 
@@ -13,47 +13,53 @@ class StatisticalMechanicsInformedLoss(LighTorchLoss):
         alpha_1: float,
         alpha_2: float,
         alpha_3: float,
-        alpha_4: float,
+        f_E: Callable,
+        f_B: Callable,
+        f_e: Callable,
+        f_p: Callable,
     ) -> None:
         super().__init__(
             labels=[
                 "Vlasov",
                 "Liouville",
-                "Quasyneutrality",
                 "Maxwell",
             ],
             factors={
                 "Vlasov": alpha_1,
                 "Liouville": alpha_2,
-                "Quasyneutrality": alpha_3,
-                "Maxwell": alpha_4,
+                "Maxwell": alpha_3,
             },
         )
         self.integral = MonteCarlo()
+        self.f_E = f_E
+        self.f_B = f_B
+        self.f_e = f_e
+        self.f_p = f_p
 
-    def vlasov(self, f: Tensor, r: Tensor, v: Tensor, t: Tensor, E: Tensor, B: Tensor) -> Tensor:
+    def vlasov(self, f_alpha: Tensor, input: Tensor, v: Tensor, E: Tensor, B: Tensor, m_alpha: float, q_alpha: float) -> Tensor:
         """
-        This is the solution for the alpha specie. The solution to the overall
+        This is the solution f_alphaor the alpha specie. The solution to the overall
         solar wind kinetic model is given by the solution to each specie separately.
         """
-        f_r = torch.autograd.grad(
-            f, r, create_graph=True
+        grad_f_alpha = torch.autograd.grad(
+            inputs = input,
+            outputs = f_alpha,
+            grad_outputs = torch.ones_like(f_alpha),
+            retain_graph=True,
+            create_graph=True
         )[0]
-        f_v = torch.autograd.grad(
[…]

### 2024-12-09 model.py

@@ -26,7 +26,7 @@ class PINN(DeepNeuralNetwork):
         )
 
     def forward(self, F: Tensor) -> Tensor:
-        return super().forward(F)
+        return super().forward()
 
 
 class Model(Module):
@@ -34,7 +34,8 @@ class Model(Module):
         super().__init__(**hparams)
         self.model = PINN(3, hparams["layers"], hparams["activations"])
         self.criterion = StatisticalMechanicsInformedLoss(
-            hparams["alpha_1"], hparams["alpha_2"]
+            hparams["alpha_1"], hparams["alpha_2"], hparams["alpha_3"],
+            PINN(), PINN(), PINN(), PINN()
         )
 
     def forward(self, x: Tensor) -> Tensor:

### 2024-12-10 paper/main.tex
default
paper

@@
 \begin{abstract}
-    %TODO
+    Solar wind modeling presents significant challenges due to its complex, quasi-neutral, and collisionless plasma dynamics.
+Traditional magnetohydrodynamic (MHD) models often fail to accurately capture the system's intricate behavior.
+This research proposes a novel approach using Statistical Mechanics Informed Neural Networks to develop a more nuanced kinetic description of solar wind phenomena.
+By integrating Vlasov-Maxwell equations with physics-informed neural network architectures, we develop a computational framework that overcomes limitations of existing analytical and numerical models.
+The proposed methodology leverages data from space observatories including DSCOVR, ACE, and WIND, establishing boundary conditions from the L1-Lagrange point to Earth's orbit.
+Our approach directly embeds physical constraints into the neural network's learning process, enabling a more interpretable and computationally efficient model.
+The research demonstrates how physics-informed machine learning can provide insights into the complex dynamics of solar wind, moving beyond traditional temperature distributions and computational constraints.By combining statistical mechanics, kinetic theory, and advanced machine learning techniques, this work offers a promising new paradigm for understanding and predicting solar wind behavior, with potential implications for space weather prediction and fundamental plasma physics research.
 \end{abstract}
@@
     \frac{\partial f_\alpha}{\partial t} + f_v \cdot \nabla_{\mathbf{r}} f_\alpha &+ \frac{q_\alpha ( f_E + f_v \times f_B)}{m_\alpha} \cdot \nabla_{\mathbf{v}} f_\alpha = 0 \text{ (Vlasov)}\\
-    \sum_{\alpha} q_\alpha \int &f_\alpha d^3(f_v) = 0 \text{ (Quasineutrality)}\\
 \frac{d}{dt} \int &f_\alpha \log f_\alpha d^3r d^3(f_v) = 0 \text{ (Liouville)}\\
-\nabla \cdot f_E &= \frac{\sum_{\alpha} \rho_\alpha}{\epsilon_0} \\
+\nabla \cdot f_E &= 0 \text{ (Quasineutrality)}\\
 \nabla \cdot f_B &= 0 \\
 \nabla \times f_E &= - \frac{\partial f_B}{\partial t} \\
-\nabla \times f_B &= \mu_0 \left(J + \epsilon_0 \frac{\partial f_E}{\partial t}\right) \text{ (Maxwell)} \\
 \end{align*}
@@
 \begin{align*}
-    r &\in \left[10R_E, 240R_E\right] \\
-    v_x &\in \left[300, 1000\right]\frac{km}{s} \\
+    r_x &\in \left[0, 230R_E\right] \\
+    v_x &\in \left[0, 1000\right]\frac{km}{s} \\
     t &\in \left[0, 7200]s \\
@@
 
-\section{Methods}
+\section{Results and discussion}
 
-\subsection{Neural Architecture Search (NAS)}
-As an intermediate step before training, a hyperparameter grid search algorithm based in Bayesian multi-objective optimization is employed to find the neural architecture.
-The Pareto front is plotted, selecting minimization of neural complexity and loss function.
+\subsection{1D-Simulations}
 
-\section{Results}
-
-\section{Discussion}
+\subsection{2D-Simulations}
 

### 2024-12-10 paper/main.tex
loss, paper

@@
     r_x &\in \left[0, 230R_E\right] \\
-    v_x &\in \left[0, 1000\right]\frac{km}{s} \\
+    v_x &\in \left[0, 1200\right]\frac{km}{s} \\
     t &\in \left[0, 7200]s \\
@@
     f_\alpha\left(r, v_{min}, t) &= 0 \\
-    f_v\left(r_{min}, t) &= \mathbf{v}_{L1}(t) \\
-    f_v\left(r_{max}, t) &= \mathbf{v}_{Earth}(t) \\
-    f_E\left(r_{min}, t\right) &= E_{L1}(t) \\
-    f_E\left(r_{max}, t\right) &= E_{Earth}(t) \\
-    f_B\left(r_{min}, t\right) &= B_{L1}(t) \\
-    f_B\left(r_{max}, t\right) &= B_{Earth}(t)
+    f_\alpha\left(r, v, t_0) &= 0 \\
+    f_E\left(r_0, t\right) &= E_{L1}(t) \\
+    f_B\left(r_0, t\right) &= B_{L1}(t) \\
 \end{align*}
@@
 
-\section{Conclusion}
+\subsection{3D-Simulations}
+
+\section{Conclusion and future work}
 This work demonstrated the effectiveness of Physics-Informed Neural Networks to model Solar Wind's behavior tied to diverse statistical mechanics formulations that lead to Vlasov's kinetic model.

### 2024-12-10 loss.py

@@ -1,11 +1,12 @@
 from lightorch.nn.criterions import LighTorchLoss
 from torch import Tensor
 from typing import Callable, Tuple
-from scipy.constants import e, m_e, m_p, mu_0, epsilon_0
+from scipy.constants import e, m_e, m_p
 import torch
 from torchquad import MonteCarlo, set_up_backend
 
-set_up_backend('torch', data_type = 'float32')
+set_up_backend("torch", data_type="float32")
+
 
 class StatisticalMechanicsInformedLoss(LighTorchLoss):
     def __init__(
@@ -36,20 +37,42 @@ class StatisticalMechanicsInformedLoss(LighTorchLoss):
         self.f_e = f_e
         self.f_p = f_p
 
-    def vlasov(self, f_alpha: Tensor, input: Tensor, v: Tensor, E: Tensor, B: Tensor, m_alpha: float, q_alpha: float) -> Tensor:
+    def boundary(self, E_L1: Tensor, B_L1: Tensor) -> Tensor:
+        f_alpha_v_min = 0.0
+        f_alpha_v_max = 0.0
+        f_alpha_t_min = 0.0
+        f_alpha_t_max = 0.0
+        f_E_r_min = E
+        f_B_r_min = B
+        return  ## revisar como hacer el dataset y demas cosas
+
+    def vlasov(
+        self,
+        f_alpha: Tensor,
+        input: Tensor,
+        v: Tensor,
+        E: Tensor,
+        B: Tensor,
+        m_alpha: float,
+        q_alpha: float,
+    ) -> Tensor:
         """
         This is the solution f_alphaor the alpha specie. The solution to the overall
         solar wind kinetic model is given by the solution to each specie separately.
         """
         grad_f_alpha = torch.autograd.grad(
-            inputs = input,
-            outputs = f_alpha,
-            grad_outputs = torch.ones_like(f_alpha),
+            inputs=input,
+            outputs=f_alpha,
+            grad_outputs=torch.ones_like(f_alpha),
             retain_graph=True,
-            create_graph=True
+            create_graph=True,
         )[0]
 
-        f_alpha_r, f_alpha_v, f_alpha_t = grad_f_alpha[:, :, :3], grad_f_alpha[:, :, 3:6], grad_f_alpha[:, :, -1] ## 3-d
+        f_alpha_r, f_alpha_v, f_alpha_t = (
+            grad_f_alpha[:, :, :3],
[…]

### 2024-12-10 model.py

@@ -2,7 +2,7 @@ from lightorch.training.supervised import Module
 from lightorch.nn import DeepNeuralNetwork
 from torch import nn, Tensor
 from .loss import StatisticalMechanicsInformedLoss
-from typing import Sequence, Union
+from typing import Sequence, Union, Dict
 
 
 class PINN(DeepNeuralNetwork):
@@ -34,9 +34,18 @@ class Model(Module):
         super().__init__(**hparams)
         self.model = PINN(3, hparams["layers"], hparams["activations"])
         self.criterion = StatisticalMechanicsInformedLoss(
-            hparams["alpha_1"], hparams["alpha_2"], hparams["alpha_3"],
-            PINN(), PINN(), PINN(), PINN()
+            hparams["alpha_1"],
+            hparams["alpha_2"],
+            hparams["alpha_3"],
+            PINN(),
+            PINN(),
+            PINN(),
+            PINN(),
         )
+
+    def loss_forward(self, batch: Tensor, idx: int) -> Dict[str, Union[Tensor, float]]:
+        input, boundary = batch
+        return dict(input=input, bc=boundary)
 
     def forward(self, x: Tensor) -> Tensor:
         return self.model(x)

### 2024-12-13 dataset.py
setting up nvidia modulus

@@ -1,0 +1,1 @@
+## define all related to the dataset

### 2024-12-13 equations.py

@@ -1,0 +1,53 @@
+from sympy import Symbol, Function
+from scipy.constants import e, m_p, m_e
+from modulus.eq.pdes import PDE
+
+
+class VlasovMaxwell(PDE):
+    def __init__(self) -> None:
+        super().__init__()
+        x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
+        t = Symbol("t")
+        input = dict(x=x, y=y, z=z, t=t)
+        v_x, v_y, v_z = Symbol("v_x"), Symbol("v_y"), Symbol("v_z")
+
+        ## Electric field
+        E_x = Function("E_x")(*input)
+        E_y = Function("E_y")(*input)
+        E_z = Function("E_z")(*input)
+
+        ## Magnetic field
+        B_x = Function("B_x")(*input)
+        B_y = Function("B_y")(*input)
+        B_z = Function("B_z")(*input)
+
+        ## Phase space descriptors for protons and electrons
+        f_e = Function(*input, v_x, v_y, v_z)
+        f_p = Function(*input, v_x, v_y, v_z)
+
+        ## PDE
+        self.equations = dict(
+            gauss_elec=E_x.diff(x, 2) + E_y.diff(y, 2) + E_z.diff(z, 2),
+            gauss_mag=B_x.diff(x, 2) + B_y.diff(y, 2) + B_z.diff(z, 2),
+            faraday_x=E_z.diff(y) - E_y.diff(z) + B_x.diff(t),
+            faraday_y=E_z.diff(x) - E_x.diff(z) + B_y.diff(t),
+            faraday_z=E_y.diff(x) - E_x.diff(y) + B_z.diff(t),
+            vlasov_proton=f_p.diff(t)
+            + (v_x * f_p.diff(x) + v_y * f_p.diff(y) + v_z * f_p.diff(z))
+            + (e / m_p)
+            * (
+                (E_x + (v_y * B_z - v_z * B_y)) * f_p.diff(v_x)
+                + (E_y + (v_x * B_z - v_z * B_x)) * f_p.diff(v_y)
+                + (E_z + (v_x * B_y - v_y * B_x)) * f_p.diff(v_z)
+            ),
+            vlasov_electron=f_e.diff(t)
+            + (v_x * f_e.diff(x) + v_y * f_e.diff(y) + v_z * f_e.diff(z))
+            + (-e / m_e)
+            * (
+                (E_x + (v_y * B_z - v_z * B_y)) * f_e.diff(v_x)
[…]

### 2024-12-13 loss.py

@@ -31,6 +31,7 @@ class StatisticalMechanicsInformedLoss(LighTorchLoss):
                 "Maxwell": alpha_3,
             },
         )
+
         self.integral = MonteCarlo()
         self.f_E = f_E
         self.f_B = f_B
@@ -60,6 +61,7 @@ class StatisticalMechanicsInformedLoss(LighTorchLoss):
         This is the solution f_alphaor the alpha specie. The solution to the overall
         solar wind kinetic model is given by the solution to each specie separately.
         """
+
         grad_f_alpha = torch.autograd.grad(
             inputs=input,
             outputs=f_alpha,
@@ -196,10 +198,12 @@ class StatisticalMechanicsInformedLoss(LighTorchLoss):
             ** 2
         )
         bc = self.boundary(*kwargs["bc"])
+
         final = (
             self.factors["Vlasov"] * vlasov
             + self.factors["Liouville"] * dS_dt
             + maxwell * self.factors["Maxwell"]
             + self.factors["BC"] * bc
         )
+
         return final, vlasov, dS_dt, maxwell

### 2024-12-13 model.py

@@ -1,51 +1,80 @@
-from lightorch.training.supervised import Module
 from lightorch.nn import DeepNeuralNetwork
-from torch import nn, Tensor
-from .loss import StatisticalMechanicsInformedLoss
-from typing import Sequence, Union, Dict
+from modulus.sym.models.arch import Arch
+from modulus.hydra import ModulusConfig
+from modulus.key import Key
+from typing import Dict
+from torch import nn
+import torch
 
 
-class PINN(DeepNeuralNetwork):
-    def __init__(
-        self,
-        in_features: int,
-        layers: Sequence[int],
-        activations: Sequence[Union[str, None]],
-    ):
+class NeuralNetwork(Arch):
+    def __init__(self, cfg: ModulusConfig):
+        if cfg.dimensions == 1:
+            self.input_keys_ = dict(
+                x=Key("x"),
+            )
+        elif cfg.dimensions == 2:
+            self.input_keys_ = dict(
+                x=Key("x"),
+                y=Key("y"),
+            )
+        elif cfg.dimensions == 3:
+            self.input_keys_ = dict(
+                x=Key("x"),
+                y=Key("y"),
+                z=Key("z"),
+            )
+        self.input_keys_.update(dict(t=Key("t")))
+
         super().__init__(
-            in_features,
-            layers,
-            list(
-                map(
-                    lambda activation: (
-                        getattr(nn, activation) if activation is not None else None
-                    ),
-                    activations,
-                )
+            input_keys=self.input_keys_,
+            output_keys=dict(
+                E=Key("E"),
+                B=Key("B"),
+                v=Key("v"),
+                f_e=Key("f_e"),
+                f_p=Key("f_p"),
             ),
         )
 
-    def forward(self, F: Tensor) -> Tensor:
-        return super().forward()
-
-
-class Model(Module):
-    def __init__(self, **hparams) -> None:
-        super().__init__(**hparams)
-        self.model = PINN(3, hparams["layers"], hparams["activations"])
-        self.criterion = StatisticalMechanicsInformedLoss(
-            hparams["alpha_1"],
-            hparams["alpha_2"],
-            hparams["alpha_3"],
-            PINN(),
-            PINN(),
-            PINN(),
-            PINN(),
+        self.f_e = DeepNeuralNetwork(
+            cfg.dimensions * 2 + 1,
[…]

### 2024-12-13 train.py

@@ -1,0 +1,77 @@
+from sympy import Symbol
+import modulus
+from modulus.hydra import ModulusConfig
+from modulus.solver import Solver
+from modulus.domain import Domain
+from modulus.domain.constraint import (
+    PointwiseBoundaryConstraint,
+    PointwiseInteriorConstraint,
+    IntegralBoundaryConstraint,
+)
+from model import NeuralNetwork
+from equations import VlasovMaxwell
+from modulus.key import Key
+
+from modulus.geometry.parametrization import Parametrization, Parameter
+from modulus.geometry.primitives_3d import Box, Sphere, Cylinder, Plane
+from modulus.geometry.primitives_2d import Rectangle, Circle, Line
+from modulus.geometry.primitives_1d import Point1D, Line1D
+from modulus.utils.io.vtk import var_to_polyvtk
+
+
+def make_geometry():
+    return ...
+
+
+@modulus.main(version_base="1.3", config_path="conf", config_name="config")
+def run(cfg: ModulusConfig) -> None:
+    x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
+    t = Symbol("t")
+    input = dict(x=x, y=y, z=z, t=t)
+    v_x, v_y, v_z = Symbol("v_x"), Symbol("v_y"), Symbol("v_z")
+
+    eq = VlasovMaxwell()
+    full_model = NeuralNetwork(cfg)
+    nodes = eq.make_nodes() + [full_model.make_node(name="neural_network")]
+    geo = make_geometry()
+    domain = Domain()
+
+    ## define L1 boundaries
+    boundary = PointwiseBoundaryConstraint(
+        nodes=nodes, geometry=geo, outvar={"E": 0}, batch_size=2
+    )
+
+    ## residual criteria
+    residual = PointwiseInteriorConstraint(
+        nodes=nodes,
+        geometry=geo,
+        outvar=dict(
+            gauss_elec=0,
+            gauss_mag=0,
+            faraday_x=0,
+            faraday_y=0,
+            faraday_z=0,
+            vlasov_proton=0,
+            vlasov_electron=0,
+        ),
+        batch_size=100,
+        bounds=dict(r=(0, 210), t=(0, 7200)),
+    )
+
+    integral = IntegralBoundaryConstraint(
+        nodes=nodes,
+        geometry=geo,
+        outvar={"VlasovMaxwell": 0},
+        batch_size=1,
+        integral_batch_size=1000,
[…]

### 2024-12-14 paper/main.tex
default

@@
 
+\subsection{Conservation of energy}
+Another important property can be provided by conservation of energy in solar wind.
+In this case, we want to account for the bulk motion of the particles; the interaction energy, this is the work of the fields; and the conservation of the pointing vector $S$ describing the energy transport throughout space and time.
+
+\begin{align*}
+    \frac{\partial}{\partial t} \left( \frac{1}{2} m_\alpha f_v^2 + m_\alpha q_\alpha f_E \cdot f_v\right) + \nabla \cdot \textbf{S} = 0
+\end{align*}
+
 \subsection{Complete dynamical system}
@@
 \nabla \times f_E &= - \frac{\partial f_B}{\partial t} \\
+\frac{\partial}{\partial t} \left( \frac{1}{2} m_\alpha f_v^2 + m_\alpha q_\alpha f_E \cdot f_v \right) &+ \nabla \cdot \textbf{S} = 0 \text{ (Conservation)} \\
 \end{align*}
@@
 
+\section{Appendix}
+
 \bibliographystyle{plain}

### 2024-12-14 equations.py
setting up boundary

@@ -1,16 +1,24 @@
-from sympy import Symbol, Function
-from scipy.constants import e, m_p, m_e
+from sympy import Symbol, Function, log
+from scipy.constants import e, m_p, m_e, mu_0
 from modulus.eq.pdes import PDE
+from itertools import chain
+from typing import List, Tuple
 
+def get_liouville(f_alpha: Function, cfg: ModulusConfig, *input):
+    integrand = f_alpha * log(f_alpha)
+    bounds: List[Tuple[float, float]] = [*chain.from_iterable(cfg.bounds)]
+    integral: int = 0
+    integral = integrand.integrate(input[0], bounds[0])
+    for x_n, bound in zip(input[1:], bounds[1:]):
+        integral = integrand.integrate(x_n, bound)
+    return integral
 
 class VlasovMaxwell(PDE):
-    def __init__(self) -> None:
+    def __init__(self, cfg: ModulusConfig) -> None:
         super().__init__()
         x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
         t = Symbol("t")
         input = dict(x=x, y=y, z=z, t=t)
-        v_x, v_y, v_z = Symbol("v_x"), Symbol("v_y"), Symbol("v_z")
-
         ## Electric field
         E_x = Function("E_x")(*input)
         E_y = Function("E_y")(*input)
@@ -21,9 +29,14 @@ class VlasovMaxwell(PDE):
         B_y = Function("B_y")(*input)
         B_z = Function("B_z")(*input)
 
+        ## Velocity field
+        V_x = Function("V_x")(*input)
+        V_y = Function("V_y")(*input)
+        V_z = Function("V_z")(*input)
+
         ## Phase space descriptors for protons and electrons
-        f_e = Function(*input, v_x, v_y, v_z)
-        f_p = Function(*input, v_x, v_y, v_z)
+        f_e = Function(*input, V_x, V_y, V_z)
+        f_p = Function(*input, V_x, V_y, V_z)
 
         ## PDE
         self.equations = dict(
@@ -33,21 +46,23 @@ class VlasovMaxwell(PDE):
             faraday_y=E_z.diff(x) - E_x.diff(z) + B_y.diff(t),
             faraday_z=E_y.diff(x) - E_x.diff(y) + B_z.diff(t),
             vlasov_proton=f_p.diff(t)
[…]

### 2024-12-14 loss.py

@@ -1,11 +1,40 @@
 from lightorch.nn.criterions import LighTorchLoss
 from torch import Tensor
-from typing import Callable, Tuple
+from typing import Callable, Tuple, Dict
 from scipy.constants import e, m_e, m_p
 import torch
 from torchquad import MonteCarlo, set_up_backend
+from modulus.loss import Loss
 
 set_up_backend("torch", data_type="float32")
+
+class LiouvilleLoss(Loss):
+    def __init__(self, cfg: ModulusConfig) -> None:
+        super().__init__()
+        self.integral_method = MonteCarlo()
+        self.dimensionality: int = cfg.dimensions * 2 + 1
+        self.sampling = cfg.sampling_monte_carlo
+        self.bounds = cfg.bounds
+
+    def forward(self, f_alpha: Callable) -> Dict[str, Tensor]:
+        entropy = self.integral.integrate(
+            lambda *args: f_alpha(*args) * torch.log(f_alpha(*args)),
+            self.dimensionality,
+            self.sampling,
+            [
+                *self.bounds.r,
+                *self.bounds.v,
+                self.bounds.t
+            ],
+        )
+
+        dS_dt = torch.autograd.grad(
+            inputs = t,
+            outputs = entropy,
+            grad_outputs = torch.ones_like(entropy),
+        )[0]
+
+        return {"Liouville": dS_dt}
 
 
 class StatisticalMechanicsInformedLoss(LighTorchLoss):

### 2024-12-14 model.py

@@ -5,33 +5,27 @@ from modulus.key import Key
 from typing import Dict
 from torch import nn
 import torch
-
+from torch import Tensor
 
 class NeuralNetwork(Arch):
     def __init__(self, cfg: ModulusConfig):
-        if cfg.dimensions == 1:
-            self.input_keys_ = dict(
-                x=Key("x"),
-            )
-        elif cfg.dimensions == 2:
-            self.input_keys_ = dict(
-                x=Key("x"),
-                y=Key("y"),
-            )
-        elif cfg.dimensions == 3:
-            self.input_keys_ = dict(
+        super().__init__(
+            input_keys = dict(
                 x=Key("x"),
                 y=Key("y"),
                 z=Key("z"),
-            )
-        self.input_keys_.update(dict(t=Key("t")))
-
-        super().__init__(
-            input_keys=self.input_keys_,
+                t=Key("t")
+            ),
             output_keys=dict(
-                E=Key("E"),
-                B=Key("B"),
-                v=Key("v"),
+                E_x=Key("E_x"),
+                E_y=Key("E_y"),
+                E_z=Key("E_z"),
+                B_x=Key("B_x"),
+                B_y=Key("B_y"),
+                B_z=Key("B_z"),
+                V_x=Key("V_x"),
+                V_y=Key("V_y"),
+                V_z=Key("V_z"),
                 f_e=Key("f_e"),
                 f_p=Key("f_p"),
             ),
@@ -67,14 +61,31 @@ class NeuralNetwork(Arch):
             [getattr(nn, f) for f in cfg.neural_network.f_B.activations],
         )
 
-    def forward(self, dict_tensor: Dict[str, torch.Tensor]):
-        input = self.concat_input(
+    def forward(self, dict_tensor: Dict[str, Tensor]):
+        input: Tensor = self.concat_input(
             {k: dict_tensor[k] for k in list(self.input_keys_.keys())},
             list(self.input_keys_.keys()),
             detach_dict=self.detach_key_dict,
             dim=-1,
         )
 
-        out = self.model(input)
+        E: Tensor = self.f_E(input)
+        B: Tensor = self.f_B(input)
+        V: Tensor = self.f_v(input)
+        f_input: Tensor = torch.cat([input, V], dim = -1)
+        f_e_output: Tensor = self.f_e(f_input, dim = -1)
[…]

### 2024-12-14 train.py

@@ -19,21 +19,19 @@ from modulus.geometry.primitives_1d import Point1D, Line1D
 from modulus.utils.io.vtk import var_to_polyvtk
 
 
-def make_geometry():
-    return ...
-
+def make_geometry(cfg: ModulusConfig):
+    nr_points: int = cfg.geometry.grid_resolution
+    ball = Sphere(center = (0, 0, 0), radius = cfg.geometry.max_length)
+    s = ball.sample_boundary(nr_points = nr_points)
+    var_to_polyvtk(s, "general_geometry")
+    return s
 
 @modulus.main(version_base="1.3", config_path="conf", config_name="config")
 def run(cfg: ModulusConfig) -> None:
-    x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
-    t = Symbol("t")
-    input = dict(x=x, y=y, z=z, t=t)
-    v_x, v_y, v_z = Symbol("v_x"), Symbol("v_y"), Symbol("v_z")
-
-    eq = VlasovMaxwell()
+    eq = VlasovMaxwell(cfg)
     full_model = NeuralNetwork(cfg)
     nodes = eq.make_nodes() + [full_model.make_node(name="neural_network")]
-    geo = make_geometry()
+    geo = make_geometry(cfg)
     domain = Domain()
 
     ## define L1 boundaries
@@ -53,20 +51,14 @@ def run(cfg: ModulusConfig) -> None:
             faraday_z=0,
             vlasov_proton=0,
             vlasov_electron=0,
+            liouville = 0,
+            energy = 0,
         ),
         batch_size=100,
-        bounds=dict(r=(0, 210), t=(0, 7200)),
+        bounds=dict(r=cfg.bounds.r, t=cfg.bounds.t),
     )
 
-    integral = IntegralBoundaryConstraint(
-        nodes=nodes,
-        geometry=geo,
-        outvar={"VlasovMaxwell": 0},
-        batch_size=1,
-        integral_batch_size=1000,
-    )  ## monte carlo integration for this one
-
-    domain.add_constraint(integral, "dirichlet")
+    domain.add_constraint(boundary, "dirichlet")
     domain.add_constraint(residual, "residual")
 
     slv = Solver(cfg, domain)

### 2024-12-15 equations.py
neural operator and training

@@ -1,4 +1,5 @@
 from sympy import Symbol, Function, log
+from modulus.hydra import ModulusConfig
 from scipy.constants import e, m_p, m_e, mu_0
 from modulus.eq.pdes import PDE
 from itertools import chain

### 2024-12-15 train.py

@@ -1,4 +1,5 @@
 from sympy import Symbol
+from argparse import ArgumentParser
 import modulus
 from modulus.hydra import ModulusConfig
 from modulus.solver import Solver
@@ -10,7 +11,6 @@ from modulus.domain.constraint import (
 )
 from model import NeuralNetwork
 from equations import VlasovMaxwell
-from modulus.key import Key
 
 from modulus.geometry.parametrization import Parametrization, Parameter
 from modulus.geometry.primitives_3d import Box, Sphere, Cylinder, Plane
@@ -26,17 +26,17 @@ def make_geometry(cfg: ModulusConfig):
     var_to_polyvtk(s, "general_geometry")
     return s
 
-@modulus.main(version_base="1.3", config_path="conf", config_name="config")
-def run(cfg: ModulusConfig) -> None:
-    eq = VlasovMaxwell(cfg)
-    full_model = NeuralNetwork(cfg)
-    nodes = eq.make_nodes() + [full_model.make_node(name="neural_network")]
+def define_constraints(cfg, nodes):
     geo = make_geometry(cfg)
     domain = Domain()
 
     ## define L1 boundaries
-    boundary = PointwiseBoundaryConstraint(
-        nodes=nodes, geometry=geo, outvar={"E": 0}, batch_size=2
+    boundary_1 = PointwiseBoundaryConstraint(
+        nodes=nodes, geometry=geo, outvar={"E": 0}, batch_size=cfg.batch_size.boundary_1
+    )
+    ## define other dirichlet
+    boundary_2 = PointwiseBoundaryConstraint(
+        nodes=nodes, geometry=geo, outvar={"E": 0}, batch_size=cfg.batch_size.boundary_1
     )
 
     ## residual criteria
@@ -54,16 +54,40 @@ def run(cfg: ModulusConfig) -> None:
             liouville = 0,
             energy = 0,
         ),
-        batch_size=100,
+        batch_size=cfg.batch_size.pde,
         bounds=dict(r=cfg.bounds.r, t=cfg.bounds.t),
     )
 
-    domain.add_constraint(boundary, "dirichlet")
+    domain.add_constraint(boundary_1, "dirichlet_1")
+    domain.add_constraint(boundary_2, "dirichlet_2")
     domain.add_constraint(residual, "residual")
+    return domain
 
+@modulus.main(version_base="1.3", config_path="conf", config_name="config")
+def train_nn(cfg: ModulusConfig) -> None:
+    eq = VlasovMaxwell(cfg)
[…]

### 2024-12-15 dataset.py
default

@@ -1,1 +1,33 @@
 ## define all related to the dataset
+import numpy as np
+from numpy.typing import NDArray
+from typing import List
+import pandas as pd
+from torch import Tensor
+import torch
+from scipy.interpolate import CubicSpline
+from torch.utils.data import Dataset
+
+def get_data_grid(real_data: NDArray, real_time: NDArray, time_grid: NDArray) -> Tensor:
+    interpolator = CubicSpline(real_time, real_data)
+    dataset = torch.from_numpy(interpolator(time_grid))
+    return dataset
+
+def get_time_grid(resolution: float) -> NDArray:
+    return np.arange(0, 1, resolution)
+
+class L1Dataset(Dataset):
+    def __init__(self, path: str, output_var: List[str]) -> None:
+        df: pd.DataFrame = pd.read_csv(path).loc[output_var, :]
+        df["t"] = np.array(list(map(lambda x: x.seconds, df.index - df.index[0])))
+        data: NDArray = df.values
+        real_data: NDArray = data
+        real_time: NDArray = data["t"]
+        self.keys = output_var
+        self.interpolator = CubicSpline(real_time, real_data)
+
+    def __getitem__(self, t):
+        values: NDArray = self.interpolator(t)
+        return {
+            key: values[idx] for idx, key in enumerate(self.keys)
+        }

### 2024-12-15 train.py

@@ -1,22 +1,22 @@
-from sympy import Symbol
+from modulus.sym.hydra import ModulusConfig
+from modulus.sym.solver import Solver
+from modulus.sym.domain import Domain
+from sympy import Symbol, Function
 from argparse import ArgumentParser
+from typing import Dict
+from torch.utils.data import Dataset
+from dataset import L1Dataset
 import modulus
-from modulus.hydra import ModulusConfig
-from modulus.solver import Solver
-from modulus.domain import Domain
-from modulus.domain.constraint import (
+from modulus.sym.domain.constraint import (
     PointwiseBoundaryConstraint,
     PointwiseInteriorConstraint,
-    IntegralBoundaryConstraint,
+    SupervisedGridConstraint
 )
-from model import NeuralNetwork
 from equations import VlasovMaxwell
+from model import NeuralNetwork
 
-from modulus.geometry.parametrization import Parametrization, Parameter
-from modulus.geometry.primitives_3d import Box, Sphere, Cylinder, Plane
-from modulus.geometry.primitives_2d import Rectangle, Circle, Line
-from modulus.geometry.primitives_1d import Point1D, Line1D
-from modulus.utils.io.vtk import var_to_polyvtk
+from modulus.sym.geometry.primitives_3d import Sphere
+from modulus.sym.utils.io.vtk import var_to_polyvtk
 
 
 def make_geometry(cfg: ModulusConfig):
@@ -27,16 +27,59 @@ def make_geometry(cfg: ModulusConfig):
     return s
 
 def define_constraints(cfg, nodes):
+    t = Symbol("t")
+    input: Dict[str, Symbol] = dict(
+        x = Symbol("x"),
+        y = Symbol("y"),
+        z = Symbol("z"),
+        t = t
+    )
+    ## Velocity field
+    V_x = Function("V_x")(*input)
+    V_y = Function("V_y")(*input)
+    V_z = Function("V_z")(*input)
     geo = make_geometry(cfg)
     domain = Domain()
+    path: str = "/data/Vlasov/curated.csv"
+    train_dataset: Dataset = L1Dataset(path, [
+        "B_x", "B_y", "B_z",
+        "E_x", "E_y", "E_z",
+        "V_x", "V_y", "V_z",
+    ])
 
     ## define L1 boundaries
-    boundary_1 = PointwiseBoundaryConstraint(
[…]

### 2024-12-15 train.py
default

@@ -27,12 +27,9 @@ def make_geometry(cfg: ModulusConfig):
     return s
 
 def define_constraints(cfg, nodes):
-    t = Symbol("t")
+    x, y, z, t = Symbol("x"), Symbol("y"), Symbol("z"), Symbol("t")
     input: Dict[str, Symbol] = dict(
-        x = Symbol("x"),
-        y = Symbol("y"),
-        z = Symbol("z"),
-        t = t
+        x = x, y = y, z = z, t = t
     )
     ## Velocity field
     V_x = Function("V_x")(*input)

### 2024-12-15 paper/main.tex
default

@@
 
-\title{Statistical Mechanics Informed Neural Networks for Solar Wind modeling}
+\title{Vlasov-Maxwell-Liouville informed Operator Learning for Solar Wind Modeling}
 \author{Jorge D. Enciso}
@@
 It uses a wide variety of model architectures (GRU, LSTM, 2D CNN ResNets, etc.) to test their effectiveness under this task.
+
+\section{Physics Informed Machine Learning}
+Usually, solely data-driven methods require huge volumes of high-quality data with black box algorithms that prevent interpretability, a key factor for physical modeling.
+The most recent research advocate for physical informed virtual loss function terms \cite{}, leading off to the vastly known field of Physics Informed Neural Networks.
+
+These models were intended to be numerical solvers for Partial Differential Equations that embed physical constraints in the loss function.
+It further extended to all fields that required PDE solving as a computationally efficient alternative to other solvers.
+
+\subsection{Physics informed Neural Networks}
+The representational power of the neural network enables learning complex non-linear functions in a given linear space.
+This representational power, coupled with suitable constraints, can lead to the major description of unknown phenomena.
+In the case of Physics informed Machine Learning, we can define a self-supervised learning procedure (governing equations), and some initial conditions, such that:
+
+\begin{align*}
+    \mathcal{L}_{total} = \mathcal{L}_{PDE} + \mathcal{L}_{Boundary}
+\end{align*}
+
+This setup lays down the foundations of Physics informed Machine Learning, where instead of relying on data-driven supervised learning, we embed governing partial differential equations that serve as a self-sufficient constraints for the training procedure, and the boundary condition defined in $\mathcal{L}_{initial}$ as a pivot.
+
+Incredibly, this framework challenges the popular believe of the reliance of deep learning architectures to high volumes of data to give meaningful results.
+Instead, we can rely on Dirichlet or Newmann boundary conditions to fulfill the initial conditions.
+
+\subsection{Approximating functionals with PINNs}
+The most recent research, using PINNs, develop a way to not just learn functions, but to learn functional operators: Lagrangian, Hamiltonian, etc.
+\cite{cranmer2020lagrangianneuralnetworks, greydanus2019hamiltonianneuralnetworks} The first research creates a neural network that takes the general coordinates $q$ and $\dot q$ as input, to provide the lagrangian as an output, taking advantage of the auto-differentiation capabilities of modern machine learning frameworks to enforce the euler-lagrange equation with gradient-descent methods:
+
+\begin{equation}
+    \frac{d}{dt} \frac{\partial \mathcal{L}}{\partial \dot q} - \frac{\partial \mathcal{L}}{\partial q} = 0
+\end{equation}
+
+The current work extrapolates this concept to the field of density functions, embedding it into the training phase by ensuring the normalized nature of a density function, the second law of thermodynamics, and systems governed by kinetic theory following Liouville's theorem.
+
+This model establishes a cornerstone for solar wind modeling, bonding the kinetic model distribution to the results of the electromagnetic model, extending the training space from     L1 Lagrange to a whole grid interval from L1 to the Earth.
+
+\subsection{Physics informed Neural Operators}
+
+\subsubsection{Fourier Neural Operator}
 
@@
 
-For the kinetic and electormagnetic model, we define the following Dirichlet boundary conditions:
+For the kinetic and electromagnetic model, we define the following Dirichlet boundary conditions:
 
@@
 
-\section{Physics Informed Machine Learning}
-Usually, solely data-driven methods require huge volumes of high-quality data with black box algorithms that prevent interpretability, a key factor for physical modeling.
-The most recent research advocate for physical informed virtual loss function terms \cite{}, leading off to the vastly known field of Physics Informed Neural Networks.
-
-These models were created as numerical methods for solving Partial Differential Equations that embed physical modeling.
-It further extended to all fields that required PDE solving as a computationally efficient numerical alternative to other solvers.
-
-\subsection{Approximating the density function with PINNs}
-The most recent research, using PINNs, develop a way to not just learn functions, but to learn functional operators: Lagrangian, Hamiltonian, etc.
-\cite{cranmer2020lagrangianneuralnetworks, greydanus2019hamiltonianneuralnetworks} The first research creates a neural network that takes the general coordinates $q$ and $\dot q$ as input, to provide the lagrangian as an output, taking advantage of the auto-differentiation capabilities of modern machine learning frameworks to enforce the euler-lagrange equation with gradient-descent methods:
-
-\begin{equation}
-    \frac{d}{dt} \frac{\partial \mathcal{L}}{\partial \dot q} - \frac{\partial \mathcal{L}}{\partial q} = 0
-\end{equation}
-
-The current work extrapolates this concept to the field of density functions, embedding it into the training phase by ensuring the normalized nature of a density function, the second law of thermodynamics, and systems governed by kinetic theory following Liouville's theorem.
-
-This model establishes a cornerstone for solar wind modeling, bonding the kinetic model distribution to the results of the electromagnetic model, extending the training space from     L1 Lagrange to a whole grid interval from L1 to the Earth.
-
-\section{Data}
-
-\subsection{DSCOVR: Deep Space Climate Observatory}
-DSCOVR, a joint mission between NASA and the National Oceanic and Atmospheric Administration, is a crucial observational platform for monitoring space weather \cite{nasa_dscovr}.
-Launched in 2015, DSCOVR's primary mission is to monitor and provide advanced warning of potentially hazardous space weather events such as solar flares and coronal mass ejections that could impact Earth.
-
-It is equipped with two key instruments for measuring both energetic particle incidence and magnetic field parameters: the Faraday cup and the magnetometer from the PlasMag instrument \cite{nasa_dscovr}.
-The readings from these two sensors are crucial for virtually analyzing plasma dynamics near the L1 Lagrange point.
-These readings will be used as part of the core model data due to their real-time availability.
-
-\subsection{ACE: Advanced Composition Explorer}
-ACE, launched in 1997, provides continuous measurements of the solar wind and interstellar particles.
-It is equipped with several instruments designed to study the composition of solar and galactic particles, which are crucial for understanding the space weather environment.
-ACE's data helps in predicting geomagnetic storms and contributes to our understanding of the heliosphere.
-
-\subsection{WIND}
-The WIND spacecraft, launched in 1994, is part of the Global Geospace Science initiative.
-It provides comprehensive measurements of the solar wind, magnetic fields, and energetic particles.
-WIND's data is essential for understanding the fundamental processes of the solar wind and its interaction with the Earth's magnetosphere.
-
 \section{Results and discussion}
 
-\subsection{1D-Simulations}
+\subsection{Neural Network 3D-Simulations}
 
-\subsection{2D-Simulations}
+\subsection{Fourier Neural Operator 3D-Simulations}
 
-\subsection{3D-Simulations}
-
-\section{Conclusion and future work}
+\section{Conclusion}
 This work demonstrated the effectiveness of Physics-Informed Neural Networks to model Solar Wind's behavior tied to diverse statistical mechanics formulations that lead to Vlasov's kinetic model.
 By embedding the governing physical equations directly into the loss function, we can efficiently train a neural network to approximate the solution of this complex, non-linear system.
-
-Future work will focus on improving the architecture design with Physics informed Neural Operators (PINOs) \cite{li2023physicsinformedneuraloperatorlearning} by embedding Maxwell's equations for data continuization through differentiable splines, the satellite data dependance inhibits the introduction os such powerful frameworks in this case.
-
-\section{Appendix}
 

### 2024-12-15 train.py

@@ -6,7 +6,7 @@ from argparse import ArgumentParser
 from typing import Dict
 from torch.utils.data import Dataset
 from dataset import L1Dataset
-import modulus
+import modulus.sym
 from modulus.sym.domain.constraint import (
     PointwiseBoundaryConstraint,
     PointwiseInteriorConstraint,
@@ -17,7 +17,6 @@ from model import NeuralNetwork
 
 from modulus.sym.geometry.primitives_3d import Sphere
 from modulus.sym.utils.io.vtk import var_to_polyvtk
-
 
 def make_geometry(cfg: ModulusConfig):
     nr_points: int = cfg.geometry.grid_resolution
@@ -116,8 +115,7 @@ def train_nn(cfg: ModulusConfig) -> None:
     slv = Solver(cfg, domain)
     slv.solve()
 
-
-@modulus.main(version_base="1.3", config_path="conf", config_name="config")
+@modulus.sym.main(version_base="1.3", config_path="conf", config_name="config")
 def train_fno(cfg: ModulusConfig) -> None:
     eq = VlasovMaxwell(cfg)
     full_model = NeuralNetwork(cfg) ## define fno
@@ -125,7 +123,6 @@ def train_fno(cfg: ModulusConfig) -> None:
     domain = define_constraints(cfg, nodes)
     slv = Solver(cfg, domain)
     slv.solve()
-
 
 if __name__ == "__main__":
     parser = ArgumentParser(prog = "VlasovMaxwell-PINN", description = "Statistical mechanics informed Neural Network training for solar wind modeling")

### 2024-12-16 data.py
setting up modulus training

@@ -1,47 +1,0 @@
-from lightning.pytorch import LightningDataModule
-from torch.utils.data import Dataset, DataLoader, random_split
-from torch import Tensor
-
-
-class dataset(Dataset):
-    def __init__(self) -> None:
-        super().__init__()
-
-    def __len__(self) -> int:
-        return None
-
-    def __getitem__(self, idx: int) -> Tensor:
-        return None
-
-
-class DataModule(LightningDataModule):
-    def __init__(
-        self, batch_size: int, num_workers: int, pin_memory: bool, train_pt: float
-    ) -> None:
-        super().__init__()
-        self.batch_size: int = batch_size
-        self.num_workers: int = num_workers
-        self.pin_memory: bool = pin_memory
-        self.train_pt: float = train_pt
-
-    def setup(self, stage: str) -> None:
-        data = dataset()
-        train_len = round(self.train_pt, (len(data)))
-        val_len = len(data) - train_len
-        self.train_ds, self.val_ds = random_split(data, [train_len, val_len])
-
-    def train_dataloader(self) -> DataLoader:
-        return DataLoader(
-            self.train_ds,
-            batch_size=self.batch_size,
-            pin_memory=self.pin_memory,
-            num_workers=self.num_workers,
-        )
-
-    def val_dataloader(self) -> DataLoader:
-        return DataLoader(
-            self.val_ds,
-            batch_size=self.batch_size,
-            pin_memory=self.pin_memory,
-            num_workers=self.num_workers,
-        )

### 2024-12-16 dataset.py

@@ -1,33 +1,20 @@
 ## define all related to the dataset
+from modulus.sym.domain.constraint.discrete import DictGridDataset
+from modulus.sym.hydra import ModulusConfig
 import numpy as np
 from numpy.typing import NDArray
-from typing import List
+from typing import List, Tuple
 import pandas as pd
 from torch import Tensor
 import torch
 from scipy.interpolate import CubicSpline
 from torch.utils.data import Dataset
 
-def get_data_grid(real_data: NDArray, real_time: NDArray, time_grid: NDArray) -> Tensor:
+
+def define_data(
+    real_data: NDArray, real_time: NDArray, cfg: ModulusConfig
+) -> Tuple[Tensor, Tensor]:
+    time_grid = np.arange(0, 1, cfg.geometry.grid_resolution)
     interpolator = CubicSpline(real_time, real_data)
     dataset = torch.from_numpy(interpolator(time_grid))
-    return dataset
-
-def get_time_grid(resolution: float) -> NDArray:
-    return np.arange(0, 1, resolution)
-
-class L1Dataset(Dataset):
-    def __init__(self, path: str, output_var: List[str]) -> None:
-        df: pd.DataFrame = pd.read_csv(path).loc[output_var, :]
-        df["t"] = np.array(list(map(lambda x: x.seconds, df.index - df.index[0])))
-        data: NDArray = df.values
-        real_data: NDArray = data
-        real_time: NDArray = data["t"]
-        self.keys = output_var
-        self.interpolator = CubicSpline(real_time, real_data)
-
-    def __getitem__(self, t):
-        values: NDArray = self.interpolator(t)
-        return {
-            key: values[idx] for idx, key in enumerate(self.keys)
-        }
+    return torch.from_numpy(time_grid), dataset

### 2024-12-16 equations.py

@@ -1,69 +1,154 @@
-from sympy import Symbol, Function, log
-from modulus.hydra import ModulusConfig
 from scipy.constants import e, m_p, m_e, mu_0
+from sympy import Symbol, Function, log
 from modulus.eq.pdes import PDE
-from itertools import chain
-from typing import List, Tuple
+from typing import Tuple
 
-def get_liouville(f_alpha: Function, cfg: ModulusConfig, *input):
-    integrand = f_alpha * log(f_alpha)
-    bounds: List[Tuple[float, float]] = [*chain.from_iterable(cfg.bounds)]
-    integral: int = 0
-    integral = integrand.integrate(input[0], bounds[0])
-    for x_n, bound in zip(input[1:], bounds[1:]):
-        integral = integrand.integrate(x_n, bound)
-    return integral
 
-class VlasovMaxwell(PDE):
-    def __init__(self, cfg: ModulusConfig) -> None:
+def define_ind() -> Tuple[Symbol, ...]:
+    x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
+    t = Symbol("t")
+    return x, y, z, t
+
+
+def define_electric_field(input) -> Tuple[Function, ...]:
+    ## Electric field
+    E_x = Function("E_x")(*input)
+    E_y = Function("E_y")(*input)
+    E_z = Function("E_z")(*input)
+    return E_x, E_y, E_z
+
+
+def define_magnetic_field(input) -> Tuple[Function, ...]:
+    ## Blectric field
+    B_x = Function("B_x")(*input)
+    B_y = Function("B_y")(*input)
+    B_z = Function("B_z")(*input)
+    return B_x, B_y, B_z
+
+
+def define_velocity_field(input) -> Tuple[Function, ...]:
+    ## Vlectric field
+    V_x = Function("V_x")(*input)
+    V_y = Function("V_y")(*input)
+    V_z = Function("V_z")(*input)
+    return V_x, V_y, V_z
+
+
+class Maxwell(PDE):
+    def __init__(self) -> None:
         super().__init__()
-        x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
-        t = Symbol("t")
+        x, y, z, t = define_ind()
         input = dict(x=x, y=y, z=z, t=t)
-        ## Electric field
-        E_x = Function("E_x")(*input)
-        E_y = Function("E_y")(*input)
-        E_z = Function("E_z")(*input)
+        B_x, B_y, B_z = define_magnetic_field(input)
[…]

### 2024-12-16 htuning.py

@@ -1,49 +1,0 @@
-from lightorch.hparams import htuning
-from typing import List, Dict
-from .data import DataModule
-from .loss import criterion
-from .model import Model
-import optuna
-
-labels: List[str] | str = criterion().labels  #
-
-
-def objective(trial: optuna.trial.Trial) -> Dict:
-    num_layers: int = trial.suggest_int("num_layers", 1, 5)
-    return dict(
-        optimizer="adam",
-        scheduler="one-cycle",
-        layers=[
-            trial.suggest_int(f"layer_{i}", 1, 128) for i in range(1, num_layers + 1)
-        ],
-        activations=[
-            trial.suggest_categorical(f"activation_{i}", ["ReLU", "Softmax"])
-            for i in range(1, num_layers + 1)
-        ],
-        lr=trial.suggest_float("lr", 1e-5, 5e-2, log=True),
-        weight_decay=trial.suggest_float("weight_decay", 1e-6, 1e-2, log=True),
-    )
-
-
-if __name__ == "__main__":
-    htuning(
-        model_class=Model,
-        hparam_objective=objective,
-        datamodule=DataModule,
-        valid_metrics=labels,
-        datamodule_kwargs=dict(pin_memory=True, num_workers=8, batch_size=32),
-        directions=["minimize" for _ in range(len(labels))],
-        precision="high",
-        n_trials=10000,
-        trainer_kwargs=dict(
-            logger=True,
-            enable_checkpointing=False,
-            max_epochs=10,
-            accelerator="cuda",
-            devices=1,
-            log_every_n_steps=22,
-            precision="32",
-            limit_train_batches=1 / 3,
-            limit_val_batches=1 / 3,
-        ),
-    )

### 2024-12-16 loss.py

@@ -1,238 +1,0 @@
-from lightorch.nn.criterions import LighTorchLoss
-from torch import Tensor
-from typing import Callable, Tuple, Dict
-from scipy.constants import e, m_e, m_p
-import torch
-from torchquad import MonteCarlo, set_up_backend
-from modulus.loss import Loss
-
-set_up_backend("torch", data_type="float32")
-
-class LiouvilleLoss(Loss):
-    def __init__(self, cfg: ModulusConfig) -> None:
-        super().__init__()
-        self.integral_method = MonteCarlo()
-        self.dimensionality: int = cfg.dimensions * 2 + 1
-        self.sampling = cfg.sampling_monte_carlo
-        self.bounds = cfg.bounds
-
-    def forward(self, f_alpha: Callable) -> Dict[str, Tensor]:
-        entropy = self.integral.integrate(
-            lambda *args: f_alpha(*args) * torch.log(f_alpha(*args)),
-            self.dimensionality,
-            self.sampling,
-            [
-                *self.bounds.r,
-                *self.bounds.v,
-                self.bounds.t
-            ],
-        )
-
-        dS_dt = torch.autograd.grad(
-            inputs = t,
-            outputs = entropy,
-            grad_outputs = torch.ones_like(entropy),
-        )[0]
-
-        return {"Liouville": dS_dt}
-
-
-class StatisticalMechanicsInformedLoss(LighTorchLoss):
-    def __init__(
-        self,
-        alpha_1: float,
-        alpha_2: float,
-        alpha_3: float,
-        f_E: Callable,
-        f_B: Callable,
-        f_e: Callable,
-        f_p: Callable,
-    ) -> None:
-        super().__init__(
-            labels=[
-                "Vlasov",
-                "Liouville",
-                "Maxwell",
-            ],
-            factors={
-                "Vlasov": alpha_1,
-                "Liouville": alpha_2,
-                "Maxwell": alpha_3,
-            },
-        )
-
-        self.integral = MonteCarlo()
-        self.f_E = f_E
-        self.f_B = f_B
-        self.f_e = f_e
-        self.f_p = f_p
-
-    def boundary(self, E_L1: Tensor, B_L1: Tensor) -> Tensor:
-        f_alpha_v_min = 0.0
-        f_alpha_v_max = 0.0
[…]

### 2024-12-16 model.py

@@ -1,56 +1,53 @@
 from lightorch.nn import DeepNeuralNetwork
 from modulus.sym.models.arch import Arch
-from modulus.hydra import ModulusConfig
-from modulus.key import Key
+from modulus.sym.models.fno.fno import FNO
+from modulus.sym.hydra import ModulusConfig
+from modulus.sym.key import Key
 from typing import Dict
 from torch import nn
 import torch
 from torch import Tensor
 
+
 class NeuralNetwork(Arch):
     def __init__(self, cfg: ModulusConfig):
         super().__init__(
-            input_keys = dict(
-                x=Key("x"),
-                y=Key("y"),
-                z=Key("z"),
-                t=Key("t")
-            ),
-            output_keys=dict(
-                E_x=Key("E_x"),
-                E_y=Key("E_y"),
-                E_z=Key("E_z"),
-                B_x=Key("B_x"),
-                B_y=Key("B_y"),
-                B_z=Key("B_z"),
-                V_x=Key("V_x"),
-                V_y=Key("V_y"),
-                V_z=Key("V_z"),
-                f_e=Key("f_e"),
-                f_p=Key("f_p"),
-            ),
+            input_keys=[Key("x"), Key("y"), Key("z"), Key("t")],
+            output_keys=[
+                Key("E_x"),
+                Key("E_y"),
+                Key("E_z"),
+                Key("B_x"),
+                Key("B_y"),
+                Key("B_z"),
+                Key("V_x"),
+                Key("V_y"),
+                Key("V_z"),
+                Key("f_e"),
+                Key("f_p"),
+            ],
         )
 
         self.f_e = DeepNeuralNetwork(
-            cfg.dimensions * 2 + 1,
+            7,
             cfg.neural_network.f_e.hidden_layers,
             [getattr(nn, f) for f in cfg.neural_network.f_e.activations],
         )
 
         self.f_p = DeepNeuralNetwork(
-            cfg.dimensions * 2 + 1,
+            7,
             cfg.neural_network.f_p.hidden_layers,
             [getattr(nn, f) for f in cfg.neural_network.f_p.activations],
         )
 
         self.f_v = DeepNeuralNetwork(
-            cfg.dimensions + 1,
+            4,
             cfg.neural_network.f_v.hidden_layers,
             [getattr(nn, f) for f in cfg.neural_network.f_v.activations],
         )
 
         self.f_E = DeepNeuralNetwork(
[…]

### 2024-12-16 train.py

@@ -1,132 +1,192 @@
 from modulus.sym.hydra import ModulusConfig
+from modulus.sym.hydra.utils import ModulusModels
 from modulus.sym.solver import Solver
 from modulus.sym.domain import Domain
-from sympy import Symbol, Function
+from sympy import Or
 from argparse import ArgumentParser
-from typing import Dict
-from torch.utils.data import Dataset
-from dataset import L1Dataset
+from typing import List, Callable
 import modulus.sym
+from modulus.sym import Node
 from modulus.sym.domain.constraint import (
     PointwiseBoundaryConstraint,
     PointwiseInteriorConstraint,
-    SupervisedGridConstraint
+    SupervisedGridConstraint,
 )
-from equations import VlasovMaxwell
-from model import NeuralNetwork
+from equations import (
+    EnergyConservation,
+    LiouvilleProton,
+    LiouvilleElectron,
+    Maxwell,
+    VlasovProton,
+    VlasovElectron,
+    define_ind,
+    define_velocity_field,
+)
+from model import NeuralNetwork, FNO
 
-from modulus.sym.geometry.primitives_3d import Sphere
+from modulus.sym.geometry.primitives_3d import Geometry, Sphere
 from modulus.sym.utils.io.vtk import var_to_polyvtk
+
 
 def make_geometry(cfg: ModulusConfig):
     nr_points: int = cfg.geometry.grid_resolution
-    ball = Sphere(center = (0, 0, 0), radius = cfg.geometry.max_length)
-    s = ball.sample_boundary(nr_points = nr_points)
+    ball = Sphere(center=(0, 0, 0), radius=cfg.geometry.max_length)
+    s = ball.sample_boundary(nr_points=nr_points)
     var_to_polyvtk(s, "general_geometry")
     return s
 
-def define_constraints(cfg, nodes):
-    x, y, z, t = Symbol("x"), Symbol("y"), Symbol("z"), Symbol("t")
-    input: Dict[str, Symbol] = dict(
-        x = x, y = y, z = z, t = t
-    )
-    ## Velocity field
-    V_x = Function("V_x")(*input)
-    V_y = Function("V_y")(*input)
-    V_z = Function("V_z")(*input)
-    geo = make_geometry(cfg)
-    domain = Domain()
-    path: str = "/data/Vlasov/curated.csv"
-    train_dataset: Dataset = L1Dataset(path, [
-        "B_x", "B_y", "B_z",
-        "E_x", "E_y", "E_z",
[…]

### 2024-12-16 training.py

@@ -1,4 +1,0 @@
-from lightorch.training.cli import trainer
-
-if __name__ == "__main__":
-    trainer(matmul_precision="high", deterministic=True, seed=123)

### 2024-12-16 train.py
default

@@ -1,5 +1,4 @@
 from modulus.sym.hydra import ModulusConfig
-from modulus.sym.hydra.utils import ModulusModels
 from modulus.sym.solver import Solver
 from modulus.sym.domain import Domain
 from sympy import Or
@@ -22,7 +21,7 @@ from equations import (
     define_ind,
     define_velocity_field,
 )
-from model import NeuralNetwork, FNO
+from model import NeuralNetwork, FNOModel
 
 from modulus.sym.geometry.primitives_3d import Geometry, Sphere
 from modulus.sym.utils.io.vtk import var_to_polyvtk
@@ -171,7 +170,7 @@ def train_nn(cfg: ModulusConfig) -> None:
 
 @modulus.sym.main(config_path="conf", config_name="config")
 def train_fno(cfg: ModulusConfig) -> None:
-    model = FNO(cfg)
+    model = FNOModel(cfg)
     nodes: List[Node] = define_nodes(model)
     geometry: Geometry = make_geometry(cfg)
     domain: Domain = Domain()

### 2024-12-16 train.py
default\

@@ -75,6 +75,15 @@ def define_pde_constraints(
             faraday_z=0,
         ),
         batch_size=cfg.batch_size.pde,
+        bounds={
+            "x": cfg.bounds.r[0],
+            "y": cfg.bounds.r[1],
+            "z": cfg.bounds.r[2],
+            "V_x": cfg.bounds.v[0],
+            "V_y": cfg.bounds.v[1],
+            "V_z": cfg.bounds.v[2],
+            "t": cfg.bounds.t,
+        },
     )
     domain.add_constraint(pde_residual, "pde_residual")
     return domain
@@ -133,9 +142,9 @@ def velocity_boundary(
         batch_size=cfg.batch_size.boundary,
         outvar={"f_e": 0, "f_p": 0},
         parameterization={
-            V_x: Or(*cfg.bounds.V_x),
-            V_y: Or(*cfg.bounds.V_y),
-            V_z: Or(*cfg.bounds.V_z),
+            V_x: Or(*cfg.bounds.v[0]),
+            V_y: Or(*cfg.bounds.v[1]),
+            V_z: Or(*cfg.bounds.v[2]),
         },
     )
     domain.add_constraint(velocity_bc, "velocity_bc")

### 2024-12-16 train.py
default

@@ -1,9 +1,12 @@
+from modulus.sym.domain.constraint.discrete import DictGridDataset
 from modulus.sym.hydra import ModulusConfig
 from modulus.sym.solver import Solver
 from modulus.sym.domain import Domain
+from numpy._typing import NDArray
+from scipy.interpolate import CubicSpline
 from sympy import Or
 from argparse import ArgumentParser
-from typing import List, Callable
+from typing import List, Callable, Tuple
 import modulus.sym
 from modulus.sym import Node
 from modulus.sym.domain.constraint import (
@@ -22,9 +25,10 @@ from equations import (
     define_velocity_field,
 )
 from model import NeuralNetwork, FNOModel
-
+import numpy as np
 from modulus.sym.geometry.primitives_3d import Geometry, Sphere
 from modulus.sym.utils.io.vtk import var_to_polyvtk
+import pandas as pd
 
 
 def make_geometry(cfg: ModulusConfig):
@@ -89,12 +93,63 @@ def define_pde_constraints(
     return domain
 
 
+def get_interpolators(variables: List[str], path: str) -> Tuple[CubicSpline, ...]:
+    interpolators: List[CubicSpline] = list()
+    data: pd.DataFrame = pd.read_csv(path)
+    time: NDArray = data.index.values.astype(np.float32)
+    for variable in variables:
+        interpolators.append(CubicSpline(time, data[variable].values))
+    return interpolators
+
+
 def l1_boundary(nodes: List[Node], domain: Domain, cfg: ModulusConfig) -> Domain:
-    train_dataset = DictGridDataset()  ## mirar como parametrizar
+    path: str = "/data/Vlasov/curated.csv"
+    variables: List[str] = [
+        "E_x",
+        "E_y",
+        "E_z",
+        "B_x",
+        "B_y",
+        "B_z",
+        "V_x",
+        "V_y",
+        "V_z",
+    ]
+    time_grid: NDArray = np.arange(0, 1, cfg.geometry.grid_resolution)
+    (
+        E_x_interpolator,
+        E_y_interpolator,
+        E_z_interpolator,
+        B_x_interpolator,
+        B_y_interpolator,
+        B_z_interpolator,
+        V_x_interpolator,
+        V_y_interpolator,
+        V_z_interpolator,
+    ) = get_interpolators(variables, path)
+    train_dataset = DictGridDataset(
+        invar=dict(
+            x=np.zeros_like(time_grid),
[…]

### 2024-12-16 dataset.py
default

@@ -14,7 +14,7 @@ from torch.utils.data import Dataset
 def define_data(
     real_data: NDArray, real_time: NDArray, cfg: ModulusConfig
 ) -> Tuple[Tensor, Tensor]:
-    time_grid = np.arange(0, 1, cfg.geometry.grid_resolution)
+    time_grid = np.arange(0, 1, cfg.custom.geometry.grid_resolution)
     interpolator = CubicSpline(real_time, real_data)
     dataset = torch.from_numpy(interpolator(time_grid))
     return torch.from_numpy(time_grid), dataset

### 2024-12-16 model.py

@@ -30,32 +30,32 @@ class NeuralNetwork(Arch):
 
         self.f_e = DeepNeuralNetwork(
             7,
-            cfg.neural_network.f_e.hidden_layers,
-            [getattr(nn, f) for f in cfg.neural_network.f_e.activations],
+            cfg.arch.neural_network.f_e.hidden_layers,
+            [getattr(nn, f) for f in cfg.arch.neural_network.f_e.activations],
         )
 
         self.f_p = DeepNeuralNetwork(
             7,
-            cfg.neural_network.f_p.hidden_layers,
-            [getattr(nn, f) for f in cfg.neural_network.f_p.activations],
+            cfg.arch.neural_network.f_p.hidden_layers,
+            [getattr(nn, f) for f in cfg.arch.neural_network.f_p.activations],
         )
 
         self.f_v = DeepNeuralNetwork(
             4,
-            cfg.neural_network.f_v.hidden_layers,
-            [getattr(nn, f) for f in cfg.neural_network.f_v.activations],
+            cfg.arch.neural_network.f_v.hidden_layers,
+            [getattr(nn, f) for f in cfg.arch.neural_network.f_v.activations],
         )
 
         self.f_E = DeepNeuralNetwork(
             4,
-            cfg.neural_network.f_E.hidden_layers,
-            [getattr(nn, f) for f in cfg.neural_network.f_E.activations],
+            cfg.arch.neural_network.f_E.hidden_layers,
+            [getattr(nn, f) for f in cfg.arch.neural_network.f_E.activations],
         )
 
         self.f_B = DeepNeuralNetwork(
-            cfg.dimensions + 1,
-            cfg.neural_network.f_B.hidden_layers,
-            [getattr(nn, f) for f in cfg.neural_network.f_B.activations],
+            4,
+            cfg.arch.neural_network.f_B.hidden_layers,
+            [getattr(nn, f) for f in cfg.arch.neural_network.f_B.activations],
         )
 
     def forward(self, dict_tensor: Dict[str, Tensor]):
@@ -117,7 +117,7 @@ class FNOModel(Arch):
                 Key("V_z"),
                 Key("t"),
             ],
-            **cfg.fno.f_e
+            **cfg.arch.fno.f_e
         )
 
         self.f_p = FNO(
@@ -130,14 +130,14 @@ class FNOModel(Arch):
[…]

### 2024-12-16 train.py

@@ -32,8 +32,8 @@ import pandas as pd
 
 
 def make_geometry(cfg: ModulusConfig):
-    nr_points: int = cfg.geometry.grid_resolution
-    ball = Sphere(center=(0, 0, 0), radius=cfg.geometry.max_length)
+    nr_points: int = cfg.custom.geometry.grid_resolution
+    ball = Sphere(center=(0, 0, 0), radius=cfg.custom.geometry.r[-1])
     s = ball.sample_boundary(nr_points=nr_points)
     var_to_polyvtk(s, "general_geometry")
     return s
@@ -80,20 +80,20 @@ def define_pde_constraints(
         ),
         batch_size=cfg.batch_size.pde,
         bounds={
-            "x": cfg.bounds.r[0],
-            "y": cfg.bounds.r[1],
-            "z": cfg.bounds.r[2],
-            "V_x": cfg.bounds.v[0],
-            "V_y": cfg.bounds.v[1],
-            "V_z": cfg.bounds.v[2],
-            "t": cfg.bounds.t,
+            "x": cfg.custom.bounds.r[0],
+            "y": cfg.custom.bounds.r[1],
+            "z": cfg.custom.bounds.r[2],
+            "V_x": cfg.custom.bounds.v[0],
+            "V_y": cfg.custom.bounds.v[1],
+            "V_z": cfg.custom.bounds.v[2],
+            "t": cfg.custom.bounds.t,
         },
     )
     domain.add_constraint(pde_residual, "pde_residual")
     return domain
 
 
-def get_interpolators(variables: List[str], path: str) -> Tuple[CubicSpline, ...]:
+def get_interpolators(variables: List[str], path: str) -> Tuple[CubicSpline]:
     interpolators: List[CubicSpline] = list()
     data: pd.DataFrame = pd.read_csv(path)
     time: NDArray = data.index.values.astype(np.float32)
@@ -115,7 +115,7 @@ def l1_boundary(nodes: List[Node], domain: Domain, cfg: ModulusConfig) -> Domain
         "V_y",
         "V_z",
     ]
-    time_grid: NDArray = np.arange(0, 1, cfg.geometry.grid_resolution)
+    time_grid: NDArray = np.arange(0, 1, cfg.custom.geometry.grid_resolution)
     (
         E_x_interpolator,
         E_y_interpolator,
@@ -135,15 +135,15 @@ def l1_boundary(nodes: List[Node], domain: Domain, cfg: ModulusConfig) -> Domain
[…]

### 2024-12-17 paper/main.tex
appendix and training

@@
 \usepackage{cite}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
 
@@
 
+Traditional Physics informed Machine Learning relies on the usage of some fixed input grid to learn a representation of the data utterly dependant on resolution.
+To solve this problem, a discretization-invariant operator should be embedded in the training process, and this is pragmatically solved with the operators/functionals (mappings between function spaces) learning.
+
+Such a general operator can be defined as a differential or integral operator.
+A useful way to embed operators, and take advantage of computational shortcuts at inference time, is using some kernel integral operator $\mathcal{G}$:
+
+\begin{align*}
+    \mathcal{G}[f(x); \kappa] = \int_\Omega \kappa(x, t) f(t) d(\alpha(t))
+\end{align*}
+
+for some region $\Omega$, defined kernel $k$, differentiable real valued functions $f$ and $\alpha$ such that the Riemann definition is fulfilled.
+
+One of such operators could be the convolution, since it is defined as follows:
+
+\begin{align*}
+    \mathcal{G}[ f(x); \kappa ] = (\kappa * f)(t) = \int_{-\infty}^\infty \kappa(\tau)f(t-\tau)\tau'
+\end{align*}
+
+The original paper presenting PINOs included 3 integral operators suitable for discretization-invariant operator learning, analogous to wavelet transforms, fourier transforms, and laplace transforms \cite{li2023physicsinformedneuraloperatorlearning}.
+
+This can be thought as a Functional analysis constraint model architecture, that pragmatically lands as a primordial Physics informed Machine Learning architecture that enables multi-resolution inference.
+
 \subsubsection{Fourier Neural Operator}
+
+To derive the definition of the Fourier Neural Operator, we must attend to the definition of the convolutional operator defined, and how it can be optimized under the convolution theorem.
+
+As it is vastly known, we can demonstrate that the following is true (Look at the appendix):
+
+\begin{align*}
+    \mathcal{G}[ f(x); \kappa ] = (\kappa * f)(t) = \int_{-\infty}^\infty \kappa(\tau)f(t-\tau)\tau' = \mathcal{F}^{-1}[\mathcal{F}[f] \cdot \mathcal{F}[\kappa]]
+\end{align*}
+
+So, we can define some frequency parameter space $\theta_\omega$, for some $R \in \theta_\omega$ such that $R = \mathcal{F}[\kappa]$, we get the following:
+
+\begin{align*}
+    \mathcal{G}[ f(x); R \in \theta_\omega] = \mathcal{F}^{-1}[\mathcal{F}[f] \cdot R]
+\end{align*}
+
+Here, we defined $\mathcal{G}$ as some operators with learnable parameters $R$ within the linear space $\theta_\omega$.
+This is the FNO, as it appropiately extrapolates some pre-defined operator to the frequency domain for computational ease.
 
@@
 Unlike ideal MHD models, which assume that the Lorentz force term can be neglected, the Vlasov equation includes this term to account for deviations from isotropy and Maxwellian distributions in the solar wind, requiring numerical approximation techniques for practical solutions \cite{grandin2023hybridvlasovsimulationsoftxray}.
-
-Note that Vlasov's equation is tightly related to statistical mechanics through Liouville's theorem:
-
-\begin{equation}
-    \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = 0
-\end{equation}
-
-If we compute use the chain rule along the phase and time space, we end up with the following expression:
-\begin{equation}
-    \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = \frac{\partial f_\alpha}{\partial t} + \nabla_{\mathbf{r}} f_\alpha \frac{\partial \mathbf{r}}{\partial t} + \nabla_{\mathbf{v}} f_\alpha \frac{\partial \mathbf{v}}{\partial t} = 0
-\end{equation}
-
-which indeed is:
-
-\begin{equation}
-    \frac{\mathrm{df_\alpha(\mathbf{r}, \mathbf{v}, t)}}{\mathrm{dt}} = \frac{\partial f_\alpha}{\partial t} + \mathbf{v} \cdot \nabla_{\mathbf{r}} f_\alpha + \mathbf{a} \cdot \nabla_{\mathbf{v}} f_\alpha = 0
-\end{equation}
 
@@
 
-A cross-training process that relies on the bonds between the Maxwell and Vlasov PDE criterions will be critical; the PDE loss APORTE to the training process will be gradually incremented through epochs, and inversely correlated with the distances from the boundary contour delimiting the training space to minimize the error margin of non-representative results in the middle of the training process.
+A cross-training process that relies on the bonds between the Maxwell and Vlasov PDE criterions will be critical; the PDE loss contribution to the training process will be gradually incremented through epochs, and inversely correlated with the distances from the boundary contour delimiting the training space to minimize the error margin of non-representative results in the middle of the training process.
 
@@
 \begin{align*}
-    f_\alpha\left(r, v_{max}, t) &= 0 \\
-    f_\alpha\left(r, v_{min}, t) &= 0 \\
-    f_\alpha\left(r, v, t_0) &= 0 \\
+    f_p(r_0, v, t) = n_{L1}(t) \left( \frac{m_p}{2 \pi k_B T_{L1}(t)}\right)^{\frac{3}{2}}& \exp\left({-\frac{m_p | v_{L1}(t) - u_{L1}(t) |^2}{2k_BT_{L1}(t)}}  \right) \\
+    f_p\left(r, v_{min}, t) &= 0 \\
+    f_p\left(r, v, t_0) &= 0 \\
+    f_e(r_0, v, t) = n_{L1}(t) \left( \frac{m_e}{2 \pi k_B T_{L1, \perp}(t)}\right)&\left( \frac{m_e}{2 \pi k_B T_{L1, \parallel}(t)}\right)^{\frac{3}{2}} \exp\left(-\frac{m_ev_{L1, \perp}^2}{2k_BT_{L1, \perp}} - \frac{m_e (v_{L1, \parallel} - u_{L1, \parallel})^2}{2k_BT_{L1, \parallel}}\right) \\
+    f_e\left(r, v_{min}, t) &= 0 \\
+    f_e\left(r, v, t_0) &= 0 \\
     f_E\left(r_0, t\right) &= E_{L1}(t) \\
+    f_E\left(r, t_0\right) &= 0 \\
     f_B\left(r_0, t\right) &= B_{L1}(t) \\
+    f_B\left(r, t_0\right) &= 0 \\
+    f_v\left(r_0, t\right) &= V_{L1}(t) \\
+    f_v\left(r, t_0\right) &= 0 \\
 \end{align*}
@@
 
+\section{Appendix}
+
+\begin{definition}[Riemann-Stieljes integral]
+    Given some partition $P = \{x_0, x_1, \ldots, x_n\}$ in $[a, b]$, we define $t_k \in [x_{k-1}, x_k]$ such that a partial sum of the form
+    \begin{align*}
+        S(P, f, \alpha) = \sum_{k = 1}^n f(t_k) \Delta \alpha_k
+    \end{align*}
+    is termed sum of Riemann-Stieljes of $f$ with respect to $\alpha$.
+$f$ is defined as Riemann-integrable with respect to $\alpha$ if $\exists N \in \mathbb{N}$ such that for every $\varepsilon > 0$ there is some $P_\varepsilon \subset [a, b]$ that encapsulates other partition $P$, and for every election of points $t_k \in [x_{k-1}, x_k]$, $\mid S(P, f, \alpha) - N \mid < \varepsilon$.
+\end{definition}
+
+\begin{definition}[Fourier Series]
+    The Fourier series of a function $f$, periodic under the interval $\Omega : [0, T]$, is given by the following series:
+    \begin{align*}
+        f(x) = \sum_{n = -\infty}^\infty c_n \exp \left( i 2\pi \frac{n}{T} x\right)
+    \end{align*}
+
+    where $c_n$ is some complex number defined within the symmetric consideration $\overline{c_k} = c_{-k}$ $ \forall k \in \mathbb{N}$ that is derived as follows:
+    \begin{align*}
+        f(x) &= \sum_{n = -\infty}^\infty c_n  e^{i 2\pi \frac{n}{T} x} \\
+        f(x) &= c_m e^{i 2\pi \frac{m}{T} x} + \sum_{n \neq m} c_n e^{i 2\pi \frac{n}{T} x} \\
+        c_m e^{i 2\pi \frac{m}{T} x} &= f(x) - \sum_{n \neq m} c_n e^{i 2\pi \frac{n}{T} x} \\
+        c_m &= e^{- i 2\pi \frac{m}{T} x} f(x) - e^{-i 2\pi \frac{m}{T} x} \sum_{n \neq m} c_n e^{i 2\pi \frac{n}{T} x} \\
+        \int_0^T c_m dx &= \int_0^Te^{- i 2\pi \frac{m}{T} x} f(x)dx  - \int_0^T \sum_{n \neq m} c_n e^{i 2\pi \frac{n - m}{T} x} dx\\
+        c_m &= \frac{1}{T} \int_0^Te^{- i 2\pi \frac{m}{T} x} f(x)dx
+    \end{align*}
+
+From this definition we can derive an interpretation of the Fourier Series as a reflection of an infinite orthogonal base with respect of the function.
+\end{definition}
+
+\begin{definition}[Fourier Transform]
+    Let $f$ be a non-periodic function. We can define the Fourier transform of $f$ as an extreme case of periodicity, interpreting it as tending to infinity:
+    \begin{align*}
+        \mathcal{F}[f] = \int_{-\infty}^{\infty} f(x) e^{- i 2 \pi s x} dx
+    \end{align*}
+\end{definition}
+
+\begin{theorem} [Convolution Theorem]
+    Given the following definition of some convolution under some domain $\Omega$ of some differentiable real valued functions $f$ and $g$:
+    \begin{align*}
+        h(x) = \int_{\Omega} f(t) g(x - t) dt
+    \end{align*}
+Given that $h \in L^1\left( \mathbb{R}^n\right)$ (the demonstration is out of scope), we can assert the following:
+\begin{align*}
+    \mathcal{F}[h] = \int_{\Omega} \left(\int_{\Omega} f(t) g(x - t) dt \right)e^{-i 2 \pi s x} dx
+\end{align*}
+Using the Fubini theorem to interchange the order of integration, we have:
+\begin{align*}
+    \mathcal{F}[h] = \int_{\Omega} \left(\int_{\Omega} f(t) g(x - t) e^{-i 2 \pi s x} dx\right) dt
+\end{align*}
+and finally, we make the following substitution $y = x - t$, with its respective jacobian being 1:
+\begin{align*}
+    \mathcal{F}[h] &= \int_{\Omega} \left(\int_{\Omega} f(t) g(y) e^{-i 2 \pi s (y + t)} dy\right) dt \\
+    \mathcal{F}[h] &= \left(\int_{\Omega} f(t) e^{-i 2 \pi s t} dt \right) \left(\int_{\Omega} g(y) e^{-i 2 \pi s y} dy \right) \\
+    \mathcal{F}[h] &= \mathcal{F}[f] \cdot \mathcal{F}[g]
+\end{align*}
+\end{theorem}
+
+\begin{definition}[Boltzmann's equation]
+    Given a probability density function $f(\mathbf{r}, \mathbf{v}, t)$ defining the probability of finding a particle in the phase space at the volume $d^3\mathbf{r}d^3\mathbf{v}$, Boltzmann's general statement is as follows:
+    \begin{align*}
+        \frac{\mathrm{df}}{\mathrm{dt}} = \left( \frac{\partial f}{\partial t} \right)_{force} + \left( \frac{\partial f}{\partial t} \right)_{coll}+ \left( \frac{\partial f}{\partial t} \right)_{diff}
+    \end{align*}
+    the $force$ term being the change due to external interference, $diff$ the diffusion of the particles, and $coll$ the forces due to collisions.
+
+    Now, given a fixed probability at time $t$ of some particles to be at $r$ and velocity $v$ in the volume $d^3\mathbf{r}d^3\mathbf{v}$, let's consider a shift due to a force $\mathbf{F}$ such that at time $t + \Delta t$, the particle is at $\mathbf{r} + \Delta \mathbf{r}= \mathbf{r} + \Delta t \mathbf{v}$ with velocity $\mathbf{v} + \Delta \mathbf{v} = \mathbf{v} + \frac{\mathbf{F}}{m} \Delta t$, and given Liouville's theorem, the phase space of probabilities doesn't change over time:
+    \begin{align*}
+        f\left(\mathbf{r} + \mathbf{v} \Delta t, \mathbf{v} + \frac{\mathbf{F}}{m} \Delta t, t + \Delta t\right) d^3\mathbf{r} d^3 \mathbf{v} = f(\mathbf{r}, \mathbf{v}, t) d^3\mathbf{r} d^3 \mathbf{v}
+    \end{align*}
+    But under collisions, we have to account for the change of the probability in the phase space due to the $\left( \frac{\partial f}{\partial t} \right)_{coll}$ term:
+    \begin{align*}
+        dN_{coll} &= \left( \frac{\partial f}{\partial t} \right)_{coll} \Delta t d^3\mathbf{r} d^3 \mathbf{v} \\
+        dN_{coll} &=  f\left(\mathbf{r} + \mathbf{v} \Delta t, \mathbf{v} + \frac{\mathbf{F}}{m} \Delta t, t + \Delta t\right) d^3\mathbf{r} d^3 \mathbf{v} - f(\mathbf{r}, \mathbf{v}, t) d^3\mathbf{r} d^3 \mathbf{v} = \Delta f d^3\mathbf{r} d^3\mathbf{v} \\
+        \frac{dN_{coll}}{d^3\mathbf{r}d^3\mathbf{v}} &= \frac{\Delta f}{\Delta t} \to^{\Delta f \to 0}_{\Delta t \to 0} \to \left(\frac{\partial f}{\partial t} \right)_{coll} = \frac{df}{dt}
+    \end{align*}
+
+Finally, if we expand upon the total differential of $f$ we get the following:
+\begin{align*}
+    df &= \frac{\partial f}{\partial t} dt + \frac{\partial f}{\partial x} dx + \frac{\partial f}{\partial y} dy + \frac{\partial f}{\partial z} dz + \frac{\partial f}{\partial v_x} dv_x + \frac{\partial f}{\partial v_y} dv_y + \frac{\partial f}{\partial v_z} dv_z \\
+    df &= \frac{\partial f}{\partial t} dt + \nabla f \cdot d\mathbf{r} + \nabla_{\mathbf{v}} f \cdot d\mathbf{v} \\
+    \frac{df}{dt} &= \frac{\partial f}{\partial t} + \nabla f \cdot \frac{d\mathbf{r}}{dt} + \nabla_{\mathbf{v}} f \cdot \frac{d\mathbf{v}}{dt} = \frac{\partial f}{\partial t} + \nabla f \cdot \mathbf{v}+ \nabla_{\mathbf{v}} f \cdot \frac{\mathbf{F}}{m}
+\end{align*}
+
+Getting our two conclusions with collisionless interaction of particles (Vlasov Equation) and collisions (Boltzmann Equation):
+\begin{align*}
+    \frac{\partial f}{\partial t} + \nabla f \cdot \mathbf{v}+ \nabla_{\mathbf{v}} f \cdot \frac{\mathbf{F}}{m} &= 0 \\
+    \frac{\partial f}{\partial t} + \nabla f \cdot \mathbf{v}+ \nabla_{\mathbf{v}} f \cdot \frac{\mathbf{F}}{m} &= \left( \frac{\partial f}{\partial t} \right)_{coll} \\
+.\end{align*}
+
+\end{definition}
+
+\begin{definition}[Lagrange multipliers]
+    Given a real valued function $f: \mathbb{R}^n \to \mathbb{R}$, let's say we want to find the local extrema of this function tied to some constraints $g_i: \mathbb{R}^n \to \mathbb{R}$.
+We define Lagrange Multipliers $\lambda_i$, real valued scalars, as follows:
+    \begin{align*}
+        \nabla f = \sum_{ i } \lambda_i \nabla g_i
+    \end{align*}
+    this definition comes from the colinearity of the gradient of constraints with respect to the function $f$. We can also create a function named \mathbf{Lagrangian function} to make the surrogate:
+    \begin{align*}
+        \mathcal{L} = f + \mathbf{\lambda} \cdot \mathbf{g}
+    \end{align*}
+    such that the following describes the extrema under defined constraints:
+    \begin{align*}
+        \nabla \mathcal{L} = 0
+    \end{align*}
+\end{definition}
+
 \bibliographystyle{plain}

### 2024-12-17 dataset.py

@@ -1,20 +1,0 @@
-## define all related to the dataset
-from modulus.sym.domain.constraint.discrete import DictGridDataset
-from modulus.sym.hydra import ModulusConfig
-import numpy as np
-from numpy.typing import NDArray
-from typing import List, Tuple
-import pandas as pd
-from torch import Tensor
-import torch
-from scipy.interpolate import CubicSpline
-from torch.utils.data import Dataset
-
-
-def define_data(
-    real_data: NDArray, real_time: NDArray, cfg: ModulusConfig
-) -> Tuple[Tensor, Tensor]:
-    time_grid = np.arange(0, 1, cfg.custom.geometry.grid_resolution)
-    interpolator = CubicSpline(real_time, real_data)
-    dataset = torch.from_numpy(interpolator(time_grid))
-    return torch.from_numpy(time_grid), dataset

### 2024-12-19 paper/main.tex
default

@@
 
-\title{Vlasov-Maxwell-Liouville informed Operator Learning for Solar Wind Modeling}
+\title{Vlasov-Maxwell informed Operator Learning for Solar Wind Modeling}
 \author{Jorge D. Enciso}
@@
 Traditional magnetohydrodynamic (MHD) models often fail to accurately capture the system's intricate behavior.
-This research proposes a novel approach using Statistical Mechanics Informed Neural Networks to develop a more nuanced kinetic description of solar wind phenomena.
-By integrating Vlasov-Maxwell equations with physics-informed neural network architectures, we develop a computational framework that overcomes limitations of existing analytical and numerical models.
+This research proposes a novel approach using Physics Informed Mahchine Learning to develop a more nuanced kinetic description of solar wind phenomena.
+By integrating Vlasov-Maxwell equations with physics-informed architectures, we develop a computational framework that overcomes limitations of existing analytical and numerical models.
 The proposed methodology leverages data from space observatories including DSCOVR, ACE, and WIND, establishing boundary conditions from the L1-Lagrange point to Earth's orbit.
 Our approach directly embeds physical constraints into the neural network's learning process, enabling a more interpretable and computationally efficient model.
-The research demonstrates how physics-informed machine learning can provide insights into the complex dynamics of solar wind, moving beyond traditional temperature distributions and computational constraints.By combining statistical mechanics, kinetic theory, and advanced machine learning techniques, this work offers a promising new paradigm for understanding and predicting solar wind behavior, with potential implications for space weather prediction and fundamental plasma physics research.
+The research demonstrates how physics-informed machine learning can provide insights into the complex dynamics of solar wind, moving beyond traditional temperature distributions and computational constraints.By combining kinetic theory and advanced machine learning techniques, this work offers a promising new paradigm for understanding and predicting solar wind behavior, with potential implications for space weather prediction and fundamental plasma physics research.
 \end{abstract}
@@
 
-\begin{equation}
-    \nabla \cdot \textbf{E} = \frac{\sigma}{\epsilon_0}
-\end{equation}
-\begin{equation}
-    \nabla \cdot \textbf{B} = 0
-\end{equation}
-\begin{equation}
-    \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0
-\end{equation}
-\begin{equation}
-    \frac{d}{dt} \left( \frac{p}{\rho^\gamma} \right)= 0
-\end{equation}
-\begin{equation}
-    \textbf{E} + \textbf{v} \times \textbf{B} = 0
-\end{equation}
-\begin{equation}
-    \frac{\partial \textbf{B}}{\partial t} = \nabla \times (\textbf{v} \times \textbf{B})
-\end{equation}
-\begin{equation}
-    \textbf{J} \times \textbf{B} = \frac{(\textbf{B}\cdot \nabla) \textbf{B}}{\mu_0} - \nabla \left(\frac{B^2}{2\mu_0}\right)
-\end{equation}
+\begin{align*}
+    \nabla \cdot \textbf{E} &= \frac{\sigma}{\epsilon_0} \\
+    \nabla \cdot \textbf{B} &= 0 \\
+    \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &= 0 \\
+    \frac{d}{dt} \left( \frac{p}{\rho^\gamma} \right) &= 0 \\
+    \textbf{E} + \textbf{v} \times \textbf{B} &= 0 \\
+    \frac{\partial \textbf{B}}{\partial t} &= \nabla \times (\textbf{v} \times \textbf{B}) \\
+    \textbf{J} \times \textbf{B} &= \frac{(\textbf{B}\cdot \nabla) \textbf{B}}{\mu_0} - \nabla \left(\frac{B^2}{2\mu_0}\right) \\
+\end{align*}
 
@@
 
-To this date, the literature doesn't embark on statistical mechanics informed neural networks for probability density function learning on a solar wind scheme.
-Therefore, the objective of this work is to leverage statistical mechanics theory and physics informed neural networks to propose a new numerical approximation for the kinetic behavior of the solar wind.
+To this date, the literature doesn't embark on physics informed neural networks for probability density function learning on a solar wind scheme.
+Therefore, the objective of this work is to leverage kinetic theory and physics informed neural networks to propose a new numerical approximation for the kinetic behavior of the solar wind.
 
@@
 \begin{align*}
-    \nabla \cdot E &= \frac{\sum_{\alpha} \rho_\alpha}{\epsilon_0} \\
+    \nabla \cdot E &= \frac{\sigma}{\epsilon_0} \\
     \nabla \cdot B &= 0 \\
@@
     \nabla \times B &= \mu_0 \left(J + \epsilon_0 \frac{\partial E}{\partial t}\right)
-\end{align*}
-
-\subsection{Conservation of energy}
-Another important property can be provided by conservation of energy in solar wind.
-In this case, we want to account for the bulk motion of the particles; the interaction energy, this is the work of the fields; and the conservation of the pointing vector $S$ describing the energy transport throughout space and time.
-
-\begin{align*}
-    \frac{\partial}{\partial t} \left( \frac{1}{2} m_\alpha f_v^2 + m_\alpha q_\alpha f_E \cdot f_v\right) + \nabla \cdot \textbf{S} = 0
 \end{align*}
@@
 \begin{align*}
-    \frac{\partial f_\alpha}{\partial t} + f_v \cdot \nabla_{\mathbf{r}} f_\alpha &+ \frac{q_\alpha ( f_E + f_v \times f_B)}{m_\alpha} \cdot \nabla_{\mathbf{v}} f_\alpha = 0 \text{ (Vlasov)}\\
-\frac{d}{dt} \int &f_\alpha \log f_\alpha d^3r d^3(f_v) = 0 \text{ (Liouville)}\\
-\nabla \cdot f_E &= 0 \text{ (Quasineutrality)}\\
+    \frac{\partial f_\alpha}{\partial t} + \mathbf{v}_\alpha \cdot \nabla_{\mathbf{r}} f_\alpha &+ \frac{q_\alpha ( f_E + \mathbf{v}_\alpha \times f_B)}{m_\alpha} \cdot \nabla_{\mathbf{v}_\alpha} f_\alpha = 0 \text{ (Vlasov)}\\
+    \frac{\partial \rho_\alpha}{\partial t} + \nabla \cdot \left(\rho_\alpha \mathbf{v}_\alpha \right) &= 0 \text{ (Continuity)}\\
+    \frac{d}{dt} \int_\Omega &f_\alpha \log f_\alpha d^3\mathbf{r} d^3\mathbf{v} = 0 \text{ (Liouville)}\\
+\nabla \cdot f_E &= \frac{\sigma_{tot}}{\epsilon_0} \approx 0 \text{ (Quasineutality)}\\
 \nabla \cdot f_B &= 0 \\
 \nabla \times f_E &= - \frac{\partial f_B}{\partial t} \\
-\frac{\partial}{\partial t} \left( \frac{1}{2} m_\alpha f_v^2 + m_\alpha q_\alpha f_E \cdot f_v \right) &+ \nabla \cdot \textbf{S} = 0 \text{ (Conservation)} \\
+\nabla \times f_B &= \mu_0 \left(\mathbf{J}_{tot} + \epsilon_0 \frac{\partial f_E}{\partial t}\right) \\
+\rho_\alpha &= \int_{\Omega_\mathbf{v}} f_\alpha d^3 \mathbf{v}_\alpha \\
+\mathbf{v}_\alpha &= \frac{1}{\rho_\alpha} \int_{\Omega_\mathbf{v}} \mathbf{v} f_\alpha d^3 \mathbf{v} \\
+\mathbf{J}_\alpha &= q_\alpha \int_{\Omega_\mathbf{v}} \mathbf{v} f_\alpha d^3\mathbf{v} = q_\alpha \mathbf{v}_\alpha \rho_\alpha\\
+\mathbf{J}_{tot} &= \sum_\alpha \mathbf{J}_\alpha
 \end{align*}
 
-$f_\alpha$ is the kinetic description of each specie, $f_E$ is the electric field descriptor, $f_B$ is the magnetic field descriptor, $f_v$ is the velocity descriptor, $J$ is the current density (derived from $f_\alpha$), and $\rho_\alpha$ is the charge density (derived from $f_\alpha$).
+$f_\alpha$ is the kinetic description of each specie, $f_E$ is the electric field descriptor, $f_B$ is the magnetic field descriptor, $f_{v, \alpha}$ is the velocity descriptor for each specie, $J$ is the current density (derived from $f_{\rho, \alpha}$), and $\rho_\alpha$ is the charge density (derived from $f_\alpha$).
 
@@
 
-For the kinetic and electromagnetic model, we define the following Dirichlet boundary conditions:
+For the kinetic and electromagnetic model, we define the following Dirichlet and high order Neumann boundary conditions:
 
 \begin{align*}
-    f_p(r_0, v, t) = n_{L1}(t) \left( \frac{m_p}{2 \pi k_B T_{L1}(t)}\right)^{\frac{3}{2}}& \exp\left({-\frac{m_p | v_{L1}(t) - u_{L1}(t) |^2}{2k_BT_{L1}(t)}}  \right) \\
+    f_p(r_0, v, t) = \rho_{p, L1}(t) &\left( \frac{m_p}{2 \pi k_B T_{p, L1}(t)}\right)^{\frac{3}{2}} \exp\left({-\frac{m_p v_{p, L1}^2}{2k_BT_{p, L1}(t)}}  \right) \\
     f_p\left(r, v_{min}, t) &= 0 \\
     f_p\left(r, v, t_0) &= 0 \\
-    f_e(r_0, v, t) = n_{L1}(t) \left( \frac{m_e}{2 \pi k_B T_{L1, \perp}(t)}\right)&\left( \frac{m_e}{2 \pi k_B T_{L1, \parallel}(t)}\right)^{\frac{3}{2}} \exp\left(-\frac{m_ev_{L1, \perp}^2}{2k_BT_{L1, \perp}} - \frac{m_e (v_{L1, \parallel} - u_{L1, \parallel})^2}{2k_BT_{L1, \parallel}}\right) \\
+        f_e(r_0, v, t) = \frac{\rho_{e, L1}(t)}{\left( 2 \pi \right)^{\frac{3}{2}} T_{e, L1, \parallel}^{\frac{1}{2}}(t)T_{e, L1, \perp}^{\frac{1}{2}}(t)} &\exp \left( - \frac{v_{e, L1, \parallel}^2}{2T_{e, L1, \parallel}} \right) \exp \left( - \frac{v_{e, L1, \perp}^2}{2T_{e, L1, \perp}} \right) \\
     f_e\left(r, v_{min}, t) &= 0 \\
     f_e\left(r, v, t_0) &= 0 \\
-    f_E\left(r_0, t\right) &= E_{L1}(t) \\
+    \nabla \cdot f_E\left(r_0, t\right) &= \frac{\rho_{L1}}{\epsilon_0}\\
     f_E\left(r, t_0\right) &= 0 \\
@@
     f_B\left(r, t_0\right) &= 0 \\
-    f_v\left(r_0, t\right) &= V_{L1}(t) \\
-    f_v\left(r, t_0\right) &= 0 \\
+    \mathbf{v}_\alpha \left(r_0, t\right) &= V_{\alpha, L1}(t) \\
 \end{align*}
+
+We use the Maxwellian and bi-Maxwellian distribution as initial conditions for $f_p$ and $f_e$ respectively, the iterative process should blend all the dynamical system components to evolve these distributions to different representations.
+
+\section{Data and preprocessing}
+This work utilizes the L1-Lagrange readings of the WIND spacecraft.
+Specifically, the Solar Wind Experiment (SWE) electron moments and multi-directional proton fluxes.
+Moreover, the 3DP PESA Low for Ion moments and the computation of the bi-Maxwellian and Maxwellian distribution at L1.
+
+To derive the parallel and perpendicular temperature of the solar wind protons with respect of the direction of the magnetic field, the following relation has been used:
+\begin{align*}
+    v_\perp &= \sqrt{\frac{2k_BT_\perp}{m_p}} \to T_\perp = \frac{v_\perp^2 m_p}{2k_B}\\
+    v_\parallel &= \sqrt{\frac{2k_BT_\parallel}{m_p}} \to T_\parallel = \frac{v_\parallel^2 m_p}{2k_B}\\
+\end{align*}
+
+$v_\perp$ and $v_\parallel$ being the thermal speeds of the protons perpendicular and parallel to the magnetic field direction respectively.
+
+The electric field components are derived from the Faraday's Cup readings as follows:
 

### 2024-12-19 download_data.py

@@ -6,14 +6,10 @@ if __name__ == "__main__":
         try:
             DataDownloading(
                 [
-                    WIND.TDP_PM(root="/data/Vlasov/TDP_PM"),
-                    DSCOVR.FaradayCup(root="/data/Vlasov/DSCOVR/FaradayCup"),
-                    DSCOVR.Magnetometer(root="/data/Vlasov/DSCOVR/Magnetometer"),
-                    ACE.SWEPAM(download_path="/data/Vlasov/ACE/SWEPAM"),
-                    ACE.SWICS(download_path="/data/Vlasov/ACE/SWICS"),
-                    ACE.EPAM(download_path="/data/Vlasov/ACE/EPAM"),
-                    ACE.MAG(download_path="/data/Vlasov/ACE/MAG"),
-                    ACE.SIS(download_path="/data/Vlasov/ACE/SIS"),
+                    WIND.SWE_electron_moments(root="/data/Vlasov/Electron"),
+                    WIND.MAG(root="/data/Vlasov/Electron"),
+                    WIND.TDP_PLSP(root="/data/Vlasov/Proton"),
+                    WIND.SWE_Ion_Anistropy(root="/data/Vlasov/Proton_anistropy"),
                 ],
                 tuple_date,
             )

### 2024-12-19 equations.py

@@ -1,7 +1,8 @@
-from scipy.constants import e, m_p, m_e, mu_0
+from scipy.constants import e, m_p, m_e, mu_0, epsilon_0
 from sympy import Symbol, Function, log
 from modulus.eq.pdes import PDE
 from typing import Tuple
+from torch import Tensor
 
 
 def define_ind() -> Tuple[Symbol, ...]:
@@ -19,19 +20,11 @@ def define_electric_field(input) -> Tuple[Function, ...]:
 
 
 def define_magnetic_field(input) -> Tuple[Function, ...]:
-    ## Blectric field
+    ## Magnetic field
     B_x = Function("B_x")(*input)
     B_y = Function("B_y")(*input)
     B_z = Function("B_z")(*input)
     return B_x, B_y, B_z
-
-
-def define_velocity_field(input) -> Tuple[Function, ...]:
-    ## Vlectric field
-    V_x = Function("V_x")(*input)
-    V_y = Function("V_y")(*input)
-    V_z = Function("V_z")(*input)
-    return V_x, V_y, V_z
 
 
 class Maxwell(PDE):
@@ -39,16 +32,51 @@ class Maxwell(PDE):
         super().__init__()
         x, y, z, t = define_ind()
         input = dict(x=x, y=y, z=z, t=t)
+
         B_x, B_y, B_z = define_magnetic_field(input)
         E_x, E_y, E_z = define_electric_field(input)
 
+        J_x = e * (n_p * v_xp - n_e * v_xe)
+        J_y = e * (n_p * v_yp - n_e * v_ye)
+        J_x = e * (n_p * v_zp - n_e * v_ze)
+
         self.equations = dict(
-            gauss_elec=E_x.diff(x, 2) + E_y.diff(y, 2) + E_z.diff(z, 2),
+            gauss_elec=E_x.diff(x, 2)
+            + E_y.diff(y, 2)
+            + E_z.diff(z, 2)
+            - (sigma / epsilon_0),
             gauss_mag=B_x.diff(x, 2) + B_y.diff(y, 2) + B_z.diff(z, 2),
             faraday_x=E_z.diff(y) - E_y.diff(z) + B_x.diff(t),
             faraday_y=E_z.diff(x) - E_x.diff(z) + B_y.diff(t),
             faraday_z=E_y.diff(x) - E_x.diff(y) + B_z.diff(t),
[…]

### 2024-12-19 model.py

@@ -1,163 +1,105 @@
+from modulus.sym.hydra import ModulusConfig
+from torch.optim.adam import Adam
+from torch.utils.data import DataLoader
 from lightorch.nn import DeepNeuralNetwork
 from modulus.sym.models.arch import Arch
-from modulus.sym.models.fno.fno import FNO
-from modulus.sym.hydra import ModulusConfig
 from modulus.sym.key import Key
-from typing import Dict
+from typing import Dict, List, Callable
 from torch import nn
-import torch
 from torch import Tensor
+from torch.optim.lr_scheduler import LambdaLR
+from modulus.models.fno import FNO
+import torch
+from equations import EnergyConservation
+from tqdm import tqdm
 
 
-class NeuralNetwork(Arch):
-    def __init__(self, cfg: ModulusConfig):
-        super().__init__(
-            input_keys=[Key("x"), Key("y"), Key("z"), Key("t")],
-            output_keys=[
-                Key("E_x"),
-                Key("E_y"),
-                Key("E_z"),
-                Key("B_x"),
-                Key("B_y"),
-                Key("B_z"),
-                Key("V_x"),
-                Key("V_y"),
-                Key("V_z"),
-                Key("f_e"),
-                Key("f_p"),
-            ],
+class PhyTrainer:
+    def __init__(
+        self,
+        epochs: int,
+        f_e: nn.Module,
+        f_p: nn.Module,
+        E: nn.Module,
+        B: nn.Module,
+        learning_rate: Dict[str, float],
+        weight_decay: Dict[str, float],
+        scheduler_stepsize: Dict[str, int],
+        lr_lambda: Dict[str, Callable[[int], float]],
+        train_loader: DataLoader,
+        val_loader: DataLoader,
+    ) -> None:
+        self.scheduler_stepsize = scheduler_stepsize
+        self.f_e = f_e
+        self.f_p = f_p
+        self.E = E
+        self.B = B
+        self.epochs: int = epochs
+        self.optimizers = dict(
+            f_e=Adam(
+                self.f_e.parameters(),
+                lr=learning_rate["f_e"],
+                weight_decay=weight_decay["f_e"],
+            ),
+            f_p=Adam(
+                self.f_p.parameters(),
+                lr=learning_rate["f_p"],
+                weight_decay=weight_decay["f_p"],
+            ),
+            E=Adam(
+                self.E.parameters(),
+                lr=learning_rate["E"],
[…]

### 2024-12-19 train.py

@@ -8,7 +8,8 @@ from sympy import Or
 from argparse import ArgumentParser
 from typing import List, Callable, Tuple
 import modulus.sym
-from modulus.sym import Node
+from modulus.sym import Node, Key
+from modulus.sym.models.fno.fno import FNO
 from modulus.sym.domain.constraint import (
     PointwiseBoundaryConstraint,
     PointwiseInteriorConstraint,
@@ -24,7 +25,7 @@ from equations import (
     define_ind,
     define_velocity_field,
 )
-from model import NeuralNetwork, FNOModel
+from model import NeuralNetwork
 import numpy as np
 from modulus.sym.geometry.primitives_3d import Geometry, Sphere
 from modulus.sym.utils.io.vtk import var_to_polyvtk
@@ -39,22 +40,29 @@ def make_geometry(cfg: ModulusConfig):
     return s
 
 
-def define_nodes(model) -> List[Node]:
+def define_pde(*args) -> List[Node]:
+    nodes = []
+    for arg in args:
+        nodes.append(arg.make_nodes())
+
     maxwell = Maxwell()
-    liouville_proton = LiouvilleProton()
-    liouville_electron = LiouvilleElectron()
-    vlasov_proton = VlasovProton()
-    vlasov_electron = VlasovElectron()
-    energy_conservation = EnergyConservation()
+
+    liouville_proton = Liouville()
+    vlasov_proton = Vlasov()
+    continuity_proton = Continuity()
+
+    liouville_electron = Liouville()
+    vlasov_electron = Vlasov()
+    continuity_electron = Continuity()
 
     nodes = (
         maxwell.nodes()
         + liouville_electron.nodes()
         + liouville_proton.nodes()
+        + continuity_proton.nodes()
         + vlasov_electron.nodes()
         + vlasov_proton.nodes()
-        + energy_conservation.nodes()
-        + [model.make_nodes()]
+        + continuity_electron.nodes()
     )
 
     return nodes
@@ -221,9 +229,9 @@ def define_boundary_conditions(
 
 
 @modulus.sym.main(config_path="conf", config_name="config")
-def train_nn(cfg: ModulusConfig) -> None:
-    model = NeuralNetwork(cfg)
-    nodes: List[Node] = define_nodes(model)
+def run(cfg: ModulusConfig) -> None:
+    f_p, f_e, f_E, f_B = ...
[…]

### 2024-12-19 equations.py
default

@@ -1,30 +1,110 @@
 from scipy.constants import e, m_p, m_e, mu_0, epsilon_0
-from sympy import Symbol, Function, log
-from modulus.eq.pdes import PDE
-from typing import Tuple
+from sympy import Symbol, Function
+from modulus.sym.eq.pdes import PDE
+from modulus.sym.geometry import Geometry
+from typing import Tuple, List
 from torch import Tensor
+from torch import nn, Tensor
+import torch
+from torchquad import MonteCarlo, set_up_backend
 
+set_up_backend('torch', data_type=torch.float32)
 
 def define_ind() -> Tuple[Symbol, ...]:
     x, y, z = Symbol("x"), Symbol("y"), Symbol("z")
     t = Symbol("t")
     return x, y, z, t
 
+def define_velocity(input, specie: str) -> Tuple[Function, ...]:
+    if specie == 'p':
+        return Function("v_xp")(*input), Function("v_yp")(*input), Function("v_zp")(*input)
+    elif specie == 'e':
+        return Function("v_xe")(*input), Function("v_ye")(*input), Function("v_ze")(*input)
+    else:
+        raise ValueError("Not valid specie")
 
 def define_electric_field(input) -> Tuple[Function, ...]:
-    ## Electric field
     E_x = Function("E_x")(*input)
     E_y = Function("E_y")(*input)
     E_z = Function("E_z")(*input)
     return E_x, E_y, E_z
 
-
 def define_magnetic_field(input) -> Tuple[Function, ...]:
-    ## Magnetic field
     B_x = Function("B_x")(*input)
     B_y = Function("B_y")(*input)
     B_z = Function("B_z")(*input)
     return B_x, B_y, B_z
+
+class Density(nn.Module):
+    def __init__(self, specie: str, geometry: Geometry, nr_points: int, montecarlo_points: int) -> None:
+        super().__init__()
+        self.specie = specie
+        self.integral = MonteCarlo()
+        self.geometry = geometry
+        self.nr_points = nr_points
+        self.monte_points = montecarlo_points
+
+    def forward(self, f_alpha: nn.Module, t: float | int) -> Tensor:
+        points = self.geometry.sample_interior(nr_points = self.nr_points)
+        output: List[Tensor] | Tensor = []
+        def integrand(rx, ry, rz, vx, vy, vz, t) -> Tensor:
+            output = f_alpha({
[…]

### 2024-12-19 model.py

@@ -1,105 +1,0 @@
-from modulus.sym.hydra import ModulusConfig
-from torch.optim.adam import Adam
-from torch.utils.data import DataLoader
-from lightorch.nn import DeepNeuralNetwork
-from modulus.sym.models.arch import Arch
-from modulus.sym.key import Key
-from typing import Dict, List, Callable
-from torch import nn
-from torch import Tensor
-from torch.optim.lr_scheduler import LambdaLR
-from modulus.models.fno import FNO
-import torch
-from equations import EnergyConservation
-from tqdm import tqdm
-
-
-class PhyTrainer:
-    def __init__(
-        self,
-        epochs: int,
-        f_e: nn.Module,
-        f_p: nn.Module,
-        E: nn.Module,
-        B: nn.Module,
-        learning_rate: Dict[str, float],
-        weight_decay: Dict[str, float],
-        scheduler_stepsize: Dict[str, int],
-        lr_lambda: Dict[str, Callable[[int], float]],
-        train_loader: DataLoader,
-        val_loader: DataLoader,
-    ) -> None:
-        self.scheduler_stepsize = scheduler_stepsize
-        self.f_e = f_e
-        self.f_p = f_p
-        self.E = E
-        self.B = B
-        self.epochs: int = epochs
-        self.optimizers = dict(
-            f_e=Adam(
-                self.f_e.parameters(),
-                lr=learning_rate["f_e"],
-                weight_decay=weight_decay["f_e"],
-            ),
-            f_p=Adam(
-                self.f_p.parameters(),
-                lr=learning_rate["f_p"],
-                weight_decay=weight_decay["f_p"],
-            ),
-            E=Adam(
-                self.E.parameters(),
-                lr=learning_rate["E"],
-                weight_decay=weight_decay["E"],
-            ),
-            B=Adam(
-                self.B.parameters(),
-                lr=learning_rate["B"],
-                weight_decay=weight_decay["B"],
-            ),
-        )
-
-        self.schedulers = dict(
-            f_e=LambdaLR(self.optimizers["f_e"], lr_lambda["f_e"]),
-            f_p=LambdaLR(
-                self.optimizers["f_p"],
-                lr_lambda["f_p"],
-            ),
-            E=LambdaLR(
-                self.optimizers["E"],
-                lr_lambda["E"],
-            ),
-            B=LambdaLR(
-                self.optimizers["B"],
-                lr_lambda["B"],
-            ),
-        )
-
[…]

### 2024-12-19 train.py

@@ -3,10 +3,8 @@ from modulus.sym.hydra import ModulusConfig
 from modulus.sym.solver import Solver
 from modulus.sym.domain import Domain
 from numpy._typing import NDArray
-from scipy.interpolate import CubicSpline
-from sympy import Or
-from argparse import ArgumentParser
-from typing import List, Callable, Tuple
+from sympy import Or, Symbol
+from typing import List, Callable
 import modulus.sym
 from modulus.sym import Node, Key
 from modulus.sym.models.fno.fno import FNO
@@ -16,16 +14,14 @@ from modulus.sym.domain.constraint import (
     SupervisedGridConstraint,
 )
 from equations import (
-    EnergyConservation,
-    LiouvilleProton,
-    LiouvilleElectron,
+    Vlasov,
+    Continuity,
     Maxwell,
-    VlasovProton,
-    VlasovElectron,
+    Velocity,
+    Density,
     define_ind,
-    define_velocity_field,
+    define_velocity
 )
-from model import NeuralNetwork
 import numpy as np
 from modulus.sym.geometry.primitives_3d import Geometry, Sphere
 from modulus.sym.utils.io.vtk import var_to_polyvtk
@@ -40,30 +36,29 @@ def make_geometry(cfg: ModulusConfig):
     return s
 
 
-def define_pde(*args) -> List[Node]:
-    nodes = []
-    for arg in args:
-        nodes.append(arg.make_nodes())
-
+def define_nodes(*args) -> List[Node]:
     maxwell = Maxwell()
-
-    liouville_proton = Liouville()
-    vlasov_proton = Vlasov()
-    continuity_proton = Continuity()
-
-    liouville_electron = Liouville()
-    vlasov_electron = Vlasov()
-    continuity_electron = Continuity()
+    vlasov_proton = Vlasov("p")
+    continuity_proton = Continuity("p")
+    vlasov_electron = Vlasov("e")
+    continuity_electron = Continuity("e")
 
     nodes = (
-        maxwell.nodes()
-        + liouville_electron.nodes()
-        + liouville_proton.nodes()
+        [
+            Node(["f_e", "t"], ["rho_e"], Density("e", geometry, nr_points, montecarlo_points)), ## mirar esto
+            Node(["f_p", "t"], ["rho_p"], Density("p", geometry, nr_points, montecarlo_points)),
[…]

### 2024-12-20 download_data.py
default

@@ -1,4 +1,4 @@
-from starstream import DataDownloading, WIND, DSCOVR, ACE
+from starstream import DataDownloading, WIND
 from utils import create_dates
 
 if __name__ == "__main__":
@@ -6,9 +6,8 @@ if __name__ == "__main__":
         try:
             DataDownloading(
                 [
-                    WIND.SWE_electron_moments(root="/data/Vlasov/Electron"),
                     WIND.MAG(root="/data/Vlasov/Electron"),
-                    WIND.TDP_PLSP(root="/data/Vlasov/Proton"),
+                    WIND.SWE_electron_moments(root="/data/Vlasov/Electron"),
                     WIND.SWE_Ion_Anistropy(root="/data/Vlasov/Proton_anistropy"),
                 ],
                 tuple_date,

### 2024-12-20 train.py

@@ -3,11 +3,12 @@ from modulus.sym.hydra import ModulusConfig
 from modulus.sym.solver import Solver
 from modulus.sym.domain import Domain
 from numpy._typing import NDArray
-from sympy import Or, Symbol
+from sympy import Symbol
 from typing import List, Callable
 import modulus.sym
 from modulus.sym import Node, Key
-from modulus.sym.models.fno.fno import FNO
+from modulus.sym.hydra import instantiate_arch
+from mudulus.sym.loss.loss import DecayedPointwiseLossNorm
 from modulus.sym.domain.constraint import (
     PointwiseBoundaryConstraint,
     PointwiseInteriorConstraint,
@@ -20,7 +21,7 @@ from equations import (
     Velocity,
     Density,
     define_ind,
-    define_velocity
+    define_velocity,
 )
 import numpy as np
 from modulus.sym.geometry.primitives_3d import Geometry, Sphere
@@ -36,7 +37,9 @@ def make_geometry(cfg: ModulusConfig):
     return s
 
 
-def define_nodes(*args) -> List[Node]:
+def define_nodes(
+    geometry: Geometry, nr_points: int, montecarlo_points: int, *args
+) -> List[Node]:
     maxwell = Maxwell()
     vlasov_proton = Vlasov("p")
     continuity_proton = Continuity("p")
@@ -45,16 +48,32 @@ def define_nodes(*args) -> List[Node]:
 
     nodes = (
         [
-            Node(["f_e", "t"], ["rho_e"], Density("e", geometry, nr_points, montecarlo_points)), ## mirar esto
-            Node(["f_p", "t"], ["rho_p"], Density("p", geometry, nr_points, montecarlo_points)),
-            Node(["f_e", "t"], ["v_xe", "v_ye", "v_ze"], Velocity("e", geometry, nr_points, montecarlo_points)),
-            Node(["f_p", "t"], ["v_xp", "v_yp", "v_zp"], Velocity("p", geometry, nr_points, montecarlo_points)),
-         ]
-        + maxwell.nodes()
-        + continuity_proton.nodes()
-        + vlasov_electron.nodes()
-        + vlasov_proton.nodes()
-        + continuity_electron.nodes()
+            Node(
+                ["f_e", "t"],
+                ["rho_e"],
+                Density("e", geometry, nr_points, montecarlo_points),
+            ),
+            Node(
[…]

### 2024-12-20 train.py
default

@@ -30,8 +30,8 @@ import pandas as pd
 
 
 def make_geometry(cfg: ModulusConfig):
-    nr_points: int = cfg.custom.geometry.grid_resolution
-    ball = Sphere(center=(0, 0, 0), radius=cfg.custom.geometry.r[-1])
+    nr_points: int = cfg.custom.geometry.nr_points
+    ball = Sphere(center=(0, 0, 0), radius=cfg.custom.bounds.r[-1])
     s = ball.sample_boundary(nr_points=nr_points)
     var_to_polyvtk(s, "general_geometry")
     return s
